### 2016-01-22 paper.tex
paper template

%% This template can be used to write a paper for
%% Computer Physics Communications using LaTeX.
%% For authors who want to write a computer program description,
%% an example Program Summary is included that only has to be
%% completed and which will give the correct layout in the
%% preprint and the journal.
%% The `elsarticle' style is used and more information on this style
%% can be found at 
%% http://www.elsevier.com/wps/find/authorsview.authors/elsarticle.
%%
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

%% This list environment is used for the references in the
%% Program Summary
%%
\newcounter{bla}
\newenvironment{refnummer}{%
\list{[\arabic{bla}]}%
{\usecounter{bla}%
 \setlength{\itemindent}{0pt}%
 \setlength{\topsep}{0pt}%
 \setlength{\itemsep}{0pt}%
 \setlength{\labelsep}{2pt}%
 \setlength{\listparindent}{0pt}%
 \settowidth{\labelwidth}{[9]}%
 \setlength{\leftmargin}{\labelwidth}%
 \addtolength{\leftmargin}{\labelsep}%
 \setlength{\rightmargin}{0pt}}}
 {\endlist}

\journal{Computer Physics Communications}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Title t.b.d}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[a]{Valerio Bertone}
\author[b]{Stefano Carrazza}
\author[a]{Nathan Hartland\corref{author}}

\cortext[author] {Corresponding author.\\\textit{E-mail address:} nathan.hartland@physics.ox.ac.uk}
\address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
\address[b]{TH Unit, CERN, CH-1211 Geneva 23, Switzerland}

\begin{abstract}
%% Text of abstract
A submitted program is expected to be of benefit to other physicists or physical chemists, or be an exemplar of good programming practice, or illustrate new or novel programming techniques which are of importance to some branch of computational physics or physical chemistry.

Acceptable program descriptions can take different forms. The following Long Write-Up structure is a suggested structure but it is not obligatory. Actual structure will depend on the length of the program, the extent to which the algorithms or software have already been described in literature, and the detail provided in the user manual.

Your manuscript and figure sources should be submitted through the Elsevier Editorial System (EES) by using the online submission tool at \\
http://www.ees.elsevier.com/cpc.

In addition to the manuscript you must supply: the program source code; job control scripts, where applicable; a README file giving the names and a brief description of all the files that make up the package and clear instructions on the installation and execution of the program; sample input and output data for at least one comprehensive test run; and, where appropriate, a user manual. These should be sent, via email as a compressed archive file, to the CPC Program Librarian at cpc@qub.ac.uk.

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
keyword1; keyword2; keyword3; etc.

\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.

{\bf PROGRAM SUMMARY/NEW VERSION PROGRAM SUMMARY}
  %Delete as appropriate.

\begin{small}
\noindent
{\em Manuscript Title:}                                       \\
{\em Authors:}                                                \\
{\em Program Title:}                                          \\
{\em Journal Reference:}                                      \\
  %Leave blank, supplied by Elsevier.
{\em Catalogue identifier:}                                   \\
  %Leave blank, supplied by Elsevier.
{\em Licensing provisions:}                                   \\
  %enter "none" if CPC non-profit use license is sufficient.
{\em Programming language:}                                   \\
{\em Computer:}                                               \\
  %Computer(s) for which program has been designed.
{\em Operating system:}                                       \\
  %Operating system(s) for which program has been designed.
{\em RAM:} bytes                                              \\
  %RAM in bytes required to execute program with typical data.
{\em Number of processors used:}                              \\
  %If more than one processor.
{\em Supplementary material:}                                 \\
  % Fill in if necessary, otherwise leave out.
{\em Keywords:} Keyword one, Keyword two, Keyword three, etc.  \\
  % Please give some freely chosen keywords that we can use in a
  % cumulative keyword index.
{\em Classification:}                                         \\
  %Classify using CPC Program Library Subject Index, see (
  % http://cpc.cs.qub.ac.uk/subjectIndex/SUBJECT_index.html)
  %e.g. 4.4 Feynman diagrams, 5 Computer Algebra.
{\em External routines/libraries:}                            \\
  % Fill in if necessary, otherwise leave out.
{\em Subprograms used:}                                       \\
  %Fill in if necessary, otherwise leave out.
{\em Catalogue identifier of previous version:}*              \\
  %Only required for a New Version summary, otherwise leave out.
{\em Journal reference of previous version:}*                  \\
  %Only required for a New Version summary, otherwise leave out.
{\em Does the new version supersede the previous version?:}*   \\
  %Only required for a New Version summary, otherwise leave out.
{\em Nature of problem:}\\
  %Describe the nature of the problem here.
   \\
{\em Solution method:}\\
  %Describe the method solution here.
   \\
{\em Reasons for the new version:}*\\
  %Only required for a New Version summary, otherwise leave out.
   \\
{\em Summary of revisions:}*\\
  %Only required for a New Version summary, otherwise leave out.
   \\
{\em Restrictions:}\\
  %Describe any restrictions on the complexity of the problem here.
   \\
{\em Unusual features:}\\
  %Describe any unusual features of the program/problem here.
   \\
{\em Additional comments:}\\
  %Provide any additional comments here.
   \\
{\em Running time:}\\
  %Give an indication of the typical running time here.
   \\

\begin{thebibliography}{0}
\bibitem{1}Reference 1         % This list should only contain those items referenced in the                 
\bibitem{2}Reference 2         % Program Summary section.   
\bibitem{3}Reference 3         % Type references in text as [1], [2], etc.
                               % This list is different from the bibliography at the end of 
                               % the Long Write-Up.
\end{thebibliography}
* Items marked with an asterisk are only required for new versions
of programs previously published in the CPC Program Library.\\
\end{small}

%% main text
\section{}
\label{}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with bibTeX database:

\bibliographystyle{elsarticle-num}
\bibliography{<your-bib-database>}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use elsarticle-num.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file

### 2016-01-25 src/APFELgrid.cc
Introduced NNPDF code to remove libnnpdf dependency

@@ -20,7 +20,7 @@
 // OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
 // SOFTWARE.
 
-#include "APFELgrid.h"
+#include "APFELgrid/APFELgrid.h"
 #include "APFEL/APFEL.h"
 
 #include "appl_grid/appl_grid.h"

### 2016-01-25 src/NNPDF/exceptions.h

@@ -1,0 +1,95 @@
+// The MIT License (MIT)
+
+// Copyright (c) 2016 Stefano Carrazza, Luigi Del Debbio, Nathan Hartland
+
+// Permission is hereby granted, free of charge, to any person obtaining a copy
+// of this software and associated documentation files (the "Software"), to deal
+// in the Software without restriction, including without limitation the rights
+// to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+// copies of the Software, and to permit persons to whom the Software is
+// furnished to do so, subject to the following conditions:
+
+// The above copyright notice and this permission notice shall be included in all
+// copies or substantial portions of the Software.
+
+// THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+// IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+// FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+// AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+// LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+// OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+// SOFTWARE.
+
+#pragma once
+
+#include <exception>
+#include <stdexcept>
+#include <string>
+
+namespace NNPDF
+{
+  /**
+   *
+   * Simple and generic exception handlers
+   *  - implements runtime and logic errors
+   *
+   */
+
+  /// Error to be thrown when runtime error is detected
+  class RuntimeException: public std::runtime_error
+  {
+  public:
+    RuntimeException(const std::string& tag, const std::string& what) : std::runtime_error(std::string("[") + std::string(tag) + std::string("] error: ") + std::string(what)) {}
+  };
+
+  /// Error to be thrown when logic error is detected
+  class LogicException: public std::logic_error
+  {
+  public:
+    LogicException(const std::string& tag, const std::string& what) : std::logic_error(std::string("[") + std::string(tag) + std::string("] error: ") + std::string(what)) {}
+  };
+
+  //_______________________________________________________________________
+  class FileError: public RuntimeException
+  {
+  public:
[…]

### 2016-01-25 src/NNPDF/fastkernel.h

@@ -1,0 +1,210 @@
+// The MIT License (MIT)
+
+// Copyright (c) Stefano Carrazza, Luigi Del Debbio, Nathan Hartland
+
+// Permission is hereby granted, free of charge, to any person obtaining a copy
+// of this software and associated documentation files (the "Software"), to deal
+// in the Software without restriction, including without limitation the rights
+// to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+// copies of the Software, and to permit persons to whom the Software is
+// furnished to do so, subject to the following conditions:
+
+// The above copyright notice and this permission notice shall be included in all
+// copies or substantial portions of the Software.
+
+// THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+// IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+// FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+// AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+// LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+// OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+// SOFTWARE.
+
+#pragma once
+
+#include <string>
+#include <vector>
+#include <istream>
+#include <fstream>
+#include <sstream>
+
+#include <map>
+
+#include "common.h"
+
+namespace NNPDF
+{
+
+    // Helper for gcc
+    template <typename T>
+    std::string ToString(T val)
+    {
+        std::stringstream stream;
+        stream << val;
+        return stream.str();
+    }
+
+   /**
+    * \class FKHeader 
+    * \brief Class for parsing the header of FK tables
+    */
+    class FKHeader
+    {
+    public:
+        FKHeader();   //!< Prototype constructor
+        FKHeader(std::istream&);                  //!< Construct from istream
+        FKHeader(std::string const& filename);   //!< Construct from file
+        FKHeader(FKHeader const&);               //!< Copy-construct
+        ~FKHeader();                             //!< Destructor
+
+        void Read(std::istream&);  //!< Read FKTable header from ostream
[…]

### 2016-01-25 src/NNPDF/fkgenerator.h

@@ -1,0 +1,64 @@
+// The MIT License (MIT)
+
+// Copyright (c) 2016 Stefano Carrazza, Luigi Del Debbio, Nathan Hartland
+
+// Permission is hereby granted, free of charge, to any person obtaining a copy
+// of this software and associated documentation files (the "Software"), to deal
+// in the Software without restriction, including without limitation the rights
+// to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+// copies of the Software, and to permit persons to whom the Software is
+// furnished to do so, subject to the following conditions:
+
+// The above copyright notice and this permission notice shall be included in all
+// copies or substantial portions of the Software.
+
+// THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+// IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+// FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+// AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+// LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+// OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+// SOFTWARE.
+
+#pragma once
+
+#include <string>
+#include <vector>
+#include "common.h"
+#include "fastkernel.h"
+
+namespace NNPDF
+{
+    /**
+    * \class FKGenerator
+    * \brief Class for filling FastKernel tables
+    */
+    class FKGenerator : public FKTable
+    {
+      private:
+        FKGenerator();                          //!< Disable default constructor
+        FKGenerator( const FKGenerator& );      //!< Disable default copy-constructor
+        FKGenerator& operator=(const FKGenerator&); //!< Disable copy-assignment
+
+      public:
+        FKGenerator( std::istream& ); // Constructor
+        virtual ~FKGenerator(); //!< Destructor
+
+         // Fill the FKTable
+        void Fill(  size_t const& d,     // Datapoint index
+                    size_t const& ix1,   // First x-index
+                    size_t const& ix2,   // Second x-index
+                    size_t const& ifl1,  // First flavour index
+                    size_t const& ifl2,  // Second flavour index
[…]

### 2016-01-25 src/fastkernel.cc

@@ -1,0 +1,886 @@
+// The MIT License (MIT)
+
+// Copyright (c) 2016 Nathan Hartland, Stefano Carrazza, Luigi Del Debbio
+
+// Permission is hereby granted, free of charge, to any person obtaining a copy
+// of this software and associated documentation files (the "Software"), to deal
+// in the Software without restriction, including without limitation the rights
+// to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+// copies of the Software, and to permit persons to whom the Software is
+// furnished to do so, subject to the following conditions:
+
+// The above copyright notice and this permission notice shall be included in all
+// copies or substantial portions of the Software.
+
+// THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+// IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+// FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+// AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+// LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+// OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+// SOFTWARE.
+
+
+#include <stdio.h>
+#include <fstream>
+#include <iostream>
+#include <ios>
+#include <iomanip>
+#include <cstdlib>
+#include <cmath>
+#include <algorithm>
+#include <sstream>
+
+#include "NNPDF/fastkernel.h"
+#include "NNPDF/exceptions.h"
+
+namespace NNPDF
+{
+
+  // Default verbosity level
+  bool FKTable::Verbose = true;
+
+  // ******************* Helpers ******************************
+
+  /**
+   * Split std::string into real std::vector
+   */
+  static void rsplit(std::vector<real>& results, std::string const& input)
+  {
+    std::stringstream strstr(input);
+    std::istream_iterator<real> it(strstr);
+    std::istream_iterator<real> end;
+    
+    results.assign(it, end);
+    return;
+  }
+
+  // *********************************************************
+
[…]

### 2016-01-25 src/fkgenerator.cc

@@ -1,0 +1,111 @@
+// The MIT License (MIT)
+
+// Copyright (c) 2016 Nathan Hartland, Stefano Carrazza, Luigi Del Debbio
+
+// Permission is hereby granted, free of charge, to any person obtaining a copy
+// of this software and associated documentation files (the "Software"), to deal
+// in the Software without restriction, including without limitation the rights
+// to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+// copies of the Software, and to permit persons to whom the Software is
+// furnished to do so, subject to the following conditions:
+
+// The above copyright notice and this permission notice shall be included in all
+// copies or substantial portions of the Software.
+
+// THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+// IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+// FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+// AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+// LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+// OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+// SOFTWARE.
+
+#include <stdio.h>
+#include <iostream>
+#include <cstdlib>
+
+#include "NNPDF/fkgenerator.h"
+#include "NNPDF/exceptions.h"
+
+namespace NNPDF
+{
+
+  FKGenerator::FKGenerator( std::istream& is ):
+  FKTable(is) 
+  {
+
+  };
+
+  FKGenerator::~FKGenerator()
+  {
+
+  }
+
+  // Fill the FKTable - beware there is no error checking performed!
+  void FKGenerator::Fill(   size_t const& d,     // Datapoint index
+                            size_t const& ix1,   // First x-index
+                            size_t const& ix2,   // Second x-index
+                            size_t const& ifl1,  // First flavour index
+                            size_t const& ifl2,  // Second flavour index
+                            real const& fk  // FK Value
+                          )
+  {
+
+    if (fk==0)
+      return;
+
[…]

### 2016-01-25 test/example.cc

@@ -13,7 +13,7 @@
 #include <cstdio>
 
 #include "APFEL/APFEL.h"
-#include "APFELgrid.h"
+#include "APFELgrid/APFELgrid.h"
 
 //#include "LHAPDF/LHAPDF.h"
 

### 2016-04-18 src/APFELgrid/fastkernel.h
AVX

@@ -462,8 +462,27 @@ namespace NNPDF
 
 
  // **********************************************************************************
+  #ifdef HAVE_AVX
+  #include <immintrin.h>
+  static inline void convolute(const float* __restrict__ x, const float* __restrict__ y, float& retval, int const& n)
+  {
+    __m256 acc = _mm256_setzero_ps();
 
-#ifdef HAVE_SSE3
+    const __m256* a = (const __m256*) x;
+    const __m256* b = (const __m256*) y;
+
+    for (int i=0; i<n/8; i++)
+      acc = _mm256_add_ps(acc, _mm256_mul_ps(*(a+i),*(b+i)));
+
+    const __m256 t1 = _mm256_hadd_ps(acc,acc);
+    const __m256 t2 = _mm256_hadd_ps(t1,t1);
+    const __m128 t3 = _mm256_extractf128_ps(t2,1);
+    const __m128 t4 = _mm_add_ss(_mm256_castps256_ps128(t2),t3);
+    retval = _mm_cvtss_f32(t4);        
+
+    _mm_empty();
+  }
+#elif defined HAVE_SSE3
   #include <pmmintrin.h>
 
   static inline void convolute(const float* __restrict__ x, const float* __restrict__ y, float& retval, int const& n)

### 2016-04-18 tests/example_conv.cc
faster SIMD with explicit _load_ps rather than casting

@@ -16,12 +16,14 @@
 #include "APFELgrid/fastkernel.h"
 #include "APFELgrid/transform.h"
 
+typedef float ctype;
+
 extern "C" void evolvepdf_(const double& , const double& , double* );
 static double* lha_pdf;
-void fkpdf (const double& x, const double& Q, const size_t& n, float* pdf)
+void fkpdf (const double& x, const double& Q, const size_t& n, ctype* pdf)
 {
   evolvepdf_(x,Q,lha_pdf);
-  NNPDF::LHA2EVLN<double, float>(lha_pdf, pdf);
+  NNPDF::LHA2EVLN<double, ctype>(lha_pdf, pdf);
 }
 
 int main(int argc, char* argv[]) {
@@ -30,15 +32,23 @@ int main(int argc, char* argv[]) {
 
 	// Read FK table
 	std::ifstream infile; infile.open("./tests/atlas-Z0-rapidity.fk");
-	NNPDF::FKTable<float> FK(infile);
+	NNPDF::FKTable<ctype> FK(infile);
 
 	// Initialise PDF set
 	LHAPDF::initPDFSet("NNPDF30_nlo_as_0118", LHAPDF::LHGRID, 0);
-	float* results = new float[FK.GetNData()];
-	FK.Convolute(fkpdf, 1, results);
+	ctype* results = new ctype[FK.GetNData()];
+
+	const int nIte = 1E5;
+	timeval startTime; gettimeofday(&startTime, NULL);
+	for (int i=0; i<nIte; i++) FK.Convolute(fkpdf, 1, results);
+	timeval endTime; gettimeofday(&endTime, NULL);
 
 	for (int i=0; i < FK.GetNData(); i++)
 		std::cout << results[i] <<std::endl;
+
+	double seconds  = endTime.tv_sec  - startTime.tv_sec;
+    double useconds = endTime.tv_usec - startTime.tv_usec;
+    std::cout << "Timer: " << (1E6f*seconds + useconds)/(nIte*FK.GetNData()) <<" us per point"<<std::endl;
 
 	delete[] results;
 	delete[] lha_pdf;

### 2016-04-18 tests/example_conv.cc
extra headers

@@ -17,9 +17,9 @@
 #include "APFELgrid/transform.h"
 
 typedef float ctype;
+static double* lha_pdf;
 
 extern "C" void evolvepdf_(const double& , const double& , double* );
-static double* lha_pdf;
 void fkpdf (const double& x, const double& Q, const size_t& n, ctype* pdf)
 {
   evolvepdf_(x,Q,lha_pdf);

### 2016-04-19 tests/example_conv.cc
notation for docco

@@ -1,55 +1,64 @@
-// example_conv.cc
-// This example demonstrates how an FK table may be convoluted with a given PDF
+// FastKernel table convolution
+// ============================
+// This example demonstrates how an **FK** table may be read from file and convoluted with a given PDF.
 
-#include <vector>
+// For this demonstration, we need some boilerplate headers
 #include <iostream>
 #include <cstdlib>
-#include <fstream>
-#include <string>
-#include <time.h>
-#include <sys/time.h>
-#include <cstdio>
 
+// Along with **LHAPDF** to provide initial scale PDFs
 #include "LHAPDF/LHAPDF.h"
 
-#include "appl_grid/appl_grid.h"
+// And the required **APFELgrid** headers
 #include "APFELgrid/fastkernel.h"
 #include "APFELgrid/transform.h"
 
+// We start with some boilerplate, a typedef named *ctype* so that we may switch between
+// double and single precision convolutions simply, and a handle for the **LHAPDF** (v5-style)
+// interface for **PDF** evolution (as per standard **APPLgrid** procedure).
 typedef float ctype;
-static double* lha_pdf;
-
 extern "C" void evolvepdf_(const double& , const double& , double* );
+
+// Now we setup an equivalent bit of boilerplate for the **FK** convolution.
+// Unlike **APPLgrids**, **FK** tables require PDFs in the DGLAP or Evolution (EVLN) basis
+// (see the reference manual for details). **APFELgrid** provides a helper utility to
+// perform the rotation to the EVLN basis from the **LHAPDF** basis. 
+// No implicit rotation is performed, as this would require prior knowledge of the
+// internal representation of PDFs used in the fit.
+
+// The function *fkpdf* therefore interfaces the **LHAPDF** call *evolvepdf_* with
+// the **FK** convolution by means of this rotation. The double *lha_pdf* holds the
+// intermediate values in the **LHAPDF** basis. Here the member argument *n* is unused.
+static double* lha_pdf;
 void fkpdf (const double& x, const double& Q, const size_t& n, ctype* pdf)
 {
   evolvepdf_(x,Q,lha_pdf);
   NNPDF::LHA2EVLN<double, ctype>(lha_pdf, pdf);
 }
 
+
+// With the boilerplate completed, we start the main loop by initialising the *lha_pdf* array
[…]

### 2016-04-19 tests/example_gen.cc

@@ -1,46 +1,77 @@
-// example_gen.cc
-// This example demonstrates how an APPLgrid can be combined
-// with evolution factors provided by APFEL to produce a FastKernel (FK)
-// table.
+// FastKernel table generation
+// ===========================
+// Here we shall demonstrate with a working example how an **APPLgrid** may be
+// combined with evolution factors provided by **APFEL** to produce a **FastKernel**
+// (**FK**) table
 
-#include <iostream>
-#include <fstream>
-#include <string>
-
+// To do this we require **APFEL**
 #include "APFEL/APFEL.h"
+// The **APFELgrid** plugin
 #include "APFELgrid/APFELgrid.h"
 #include "APFELgrid/fastkernel.h"
-
+// and **APPLgrid**
 #include "appl_grid/appl_grid.h"
 
-static const double Q0 = 1.0; // 1 GeV initial scale
+// We will also demonstrate how said **FK** table may be written to and read from file.
+// To do so we will need the following headers.
+#include <iostream>
+#include <fstream>
 
+// Beginning the main loop
 int main(int argc, char* argv[]) {
 
-  // Initialise APFEL
-  APFEL::SetTheory(string("QCD"));
-  APFEL::SetAlphaQCDRef(0.118, 91.2);
-  APFEL::SetPerturbativeOrder(1);
-
+  // We start by initialising the general evolution parameters of **APFEL**.
+  // Here we are using the exact solution of NLO DGLAP with a strong coupling
+  // of 0.118 evaluated at the Z mass. Our evolution and alpha_S running is limited
+  // to five flavours
+  APFEL::SetTheory("QCD");
   APFEL::SetPDFEvolution("exactalpha");
   APFEL::SetAlphaEvolution("exact");
 
-  // Needs to match with the APPLgrid
+  APFEL::SetAlphaQCDRef(0.118, 91.2);
+  APFEL::SetPerturbativeOrder(1);
   APFEL::SetMaxFlavourAlpha(5);
   APFEL::SetMaxFlavourPDFs(5);
 
-  // Read APPLgrid
+  // Next we read the **APPLgrid** through the usual procedure
   appl::grid g("./tests/atlas-Z0-rapidity.root");
 
-  // Set x-grid
+  // Once the **APPLgrid** is read, we can specify the initial-scale grid of x-points
+  // to be used for the interpolation of the evolution factors. To do this
+  // we make use of the **APFELgrid** helper function *get_appl_Xmin* which returns the lowest
[…]

### 2016-04-19 docco_generate.sh
docs

@@ -1,0 +1,1 @@
+docco -L docco_cpp.json -l linear ./README.md ./tests/example_gen.cc ./tests/example_conv.cc

### 2016-04-19 tests/example_conv.cc

@@ -1,8 +1,10 @@
+// APFELgrid
+// =========
 // FastKernel table convolution
-// ============================
+// ----------------------------
 // This example demonstrates how an **FK** table may be read from file and convoluted with a given PDF.
 
-// For this demonstration, we need some boilerplate headers
+// For this demonstration, we need some standard headers
 #include <iostream>
 #include <cstdlib>
 

### 2016-04-19 tests/example_gen.cc

@@ -1,5 +1,7 @@
+// APFELgrid
+// =========
 // FastKernel table generation
-// ===========================
+// ---------------------------
 // Here we shall demonstrate with a working example how an **APPLgrid** may be
 // combined with evolution factors provided by **APFEL** to produce a **FastKernel**
 // (**FK**) table

### 2016-04-19 paper.tex
first version of the paper: abstract + introduction

@@
 
-\title{Title t.b.d}
+\title{{\tt APFELgrid}: a fast interface to hadronic observables for
+  fits of parton densities}
 
@@
 \author[b]{Stefano Carrazza}
-\author[a]{Nathan Hartland\corref{author}}
+\author[a]{Nathan P. Hartland\corref{author}}
 
@@
 \begin{abstract}
-%% Text of abstract
-A submitted program is expected to be of benefit to other physicists or physical chemists, or be an exemplar of good programming practice, or illustrate new or novel programming techniques which are of importance to some branch of computational physics or physical chemistry.
-
-Acceptable program descriptions can take different forms.
-The following Long Write-Up structure is a suggested structure but it is not obligatory.
-Actual structure will depend on the length of the program, the extent to which the algorithms or software have already been described in literature, and the detail provided in the user manual.
-
-Your manuscript and figure sources should be submitted through the Elsevier Editorial System (EES) by using the online submission tool at \\
-http://www.ees.elsevier.com/cpc.
-
-In addition to the manuscript you must supply: the program source code; job control scripts, where applicable; a README file giving the names and a brief description of all the files that make up the package and clear instructions on the installation and execution of the program; sample input and output data for at least one comprehensive test run; and, where appropriate, a user manual.
-These should be sent, via email as a compressed archive file, to the CPC Program Librarian at cpc@qub.ac.uk.
-
+We present a new method conceived to ease the inclusion of hadronic
+observables into parton distribution function (PDF) fits. Our
+technique relies on the the same principles adopted by the existing
+fast interfaces but implements a set of improvements aimed at
+optimizing the performance in the context of PDF fits where many
+iterations and many predictions for each iteration are usually
+required. The main novelties are the precomputation of the PDF and
+$\alpha_s$ evolution and the optimization of the numerical convolution
+between hard cross sections and PDFs. We demostrate that our techique
+can lead to a substantial speed improvement as compared to the
+existing interfaces.
 \end{abstract}
@@
 %% keywords here, in the form: keyword \sep keyword
-keyword1; keyword2; keyword3; etc.
+QCD; parton distribution functions; fast predictions.
 
@@
 
-%%
-%% Start line numbering here if you want
-%%
-% \linenumbers
-
-% Computer program descriptions should contain the following
-% PROGRAM SUMMARY.
-
-{\bf PROGRAM SUMMARY/NEW VERSION PROGRAM SUMMARY}
-  %Delete as appropriate.
-
 \begin{small}
 \noindent
-{\em Manuscript Title:}                                       \\
-{\em Authors:}                                                \\
-{\em Program Title:}                                          \\
+{\em Manuscript Title:} {\tt APFELgrid}: a fast interface to hadronic observables for
+  fits of parton densities                                      \\
+{\em Authors:} V.~Bertone, S.~Carrazza, N.P.~Hartland                                               \\
+{\em Program Title:} {\tt APFELgrid}                                          \\
 {\em Journal Reference:}                                      \\
@@
    \\
-
-\begin{thebibliography}{0}
-\bibitem{1}Reference 1         % This list should only contain those items referenced in the                 
-\bibitem{2}Reference 2         % Program Summary section.   
-\bibitem{3}Reference 3         % Type references in text as [1], [2], etc.
-                               % This list is different from the bibliography at the end of 
-                               % the Long Write-Up.
-\end{thebibliography}
 * Items marked with an asterisk are only required for new versions
@@
 %% main text
-\section{}
-\label{}
+\section{Introduction}\label{sec:intro}
+
+The need for fast and accurate cross-section predictions in the Large
+Hadron Collider (LHC) era is becoming increasingly important. In fact,
+the outstanding amount of precise data produced by the LHC experiments
+is bound to provide a very stringent constraint on the proton parton
+distribution functions (PDFs) on the condition that as many
+measurements as possible are included in PDF fits with the highest
+accuracy possible. However, the direct use of most of the codes that
+provide predictions for non-inclusive hadron-collider observables in a
+PDF fit, which typically requires a large number of iterations, is not
+viable because the time required to obtain accurate predictions is
+usually of the order of a few hours or more. In order to overcome this
+limitation, the typical strategy adopted relies on the precomputation
+of the partonic hard cross sections in such a way that the numerical
+convolution with any set of PDFs can be quickly performed {\it a
+  posteriori} by means of interpolation techniques. Examples of the
+implementation of such a strategy are {\tt
+  APPLgrid}~\cite{Carli:2010rw} and {\tt
+  FastNLO}~\cite{Wobisch:2011ij}.
 
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{<your-bib-database>}
+\bibliography{paper.bib}
 

### 2016-04-20 src/APFELgrid.cc
removed dependence on pdfweight parameter

@@ -26,6 +26,9 @@
 #include "appl_grid/appl_grid.h"
 #include "appl_grid/appl_igrid.h"
 
+#include "TFile.h"
+#include "TVectorT.h"
+
 #include <math.h>
 #include <sys/time.h>
 
@@ -170,6 +173,13 @@ namespace APFELgrid{
     if (g.calculation() == appl::grid::AMCATNLO)
       return (pto==0) ? 3:0;
     return pto;
+  }
+
+  // Fetches the flag specifying whether or not a grid is weighted from the TFile
+  bool get_pdf_wgt(TFile& f, std::string const& directory,  int const& gidx, int const& d)
+  {
+    char name[128];  sprintf(name, (directory+"/weight[alpha-%d][%03d]").c_str(), gidx, d); 
+    std::string strname = name; return (*(TVectorT<double>*)f.Get((strname+"/Parameters").c_str()))(13);
   }
 
   // Returns the APPLgrid PDF object associated with the ith subgrid of g
@@ -355,11 +365,14 @@ namespace APFELgrid{
   }
 
   // Performs the combination of an APPLgrid g with evolution factors provided
-  // by APFEL, resulting in a new FK table. Required arguments are the name of the produced table 'name', 
-  // the appl::grid g, the initial scale for the FK tables Q0, and a boolean specifying whether or not the
-  // APPLgrid has PDF weights enabled (this is at the moment impossible to tell from the APPLgrid API) 
-  NNPDF::FKTable<double>* computeFK( std::string const& name, appl::grid const& g, double const& Q0, bool const& pdfwgt )
+  // by APFEL, resulting in a new FK table. Required arguments are the initial scale Q0, name of the produced table 'name', 
+  // the appl::grid g, the path to the appl::grid file itself, and an (optional) appl::grid directory. The paths are required
+  // as we have to reconstruct the _m_reweight parameter from APPLgrid.  
+  NNPDF::FKTable<double>* computeFK( double const& Q0, std::string const& name, appl::grid const& g, std::string const& gridfile, std::string directory)
   {
+    // Read TFile for extraction of pdfwgt parameter
+    TFile f(gridfile.c_str());
+
     // Set APFEL scale limits
     double Qmin, Qmax;
     get_appl_Q2lims(g, Qmin, Qmax);
@@ -388,6 +401,7 @@ namespace APFELgrid{
     {
[…]

### 2016-04-20 src/APFELgrid/APFELgrid.h

@@ -42,10 +42,10 @@ namespace APFELgrid
 
   // ************************ FK table computation **************************
   // Performs the combination of an APPLgrid g with evolution factors provided
-  // by APFEL, resulting in a new FK table. Required arguments are the name of the produced table 'name', 
-  // the appl::grid g, the initial scale for the FK tables Q0, and a boolean specifying whether or not the
-  // APPLgrid has PDF weights enabled (this is at the moment impossible to tell from the APPLgrid API) 
-  NNPDF::FKTable<double>* computeFK( std::string const& name, appl::grid const& g, double const& Q0, bool const& pdfwgt );
+  // by APFEL, resulting in a new FK table. Required arguments are the initial 
+  // scale for the FK tables (Q0), the name of the produced table (name), the appl::grid (g),
+  // the path to the appl::grid (gridfile) and an optional appl::grid directory (directory).
+  NNPDF::FKTable<double>* computeFK( double const& Q0, std::string const& name, appl::grid const& g, std::string const& gridfile, std::string directory = "grid" );
 }
 
 namespace NNPDF

### 2016-04-20 tests/example_gen.cc

@@ -36,7 +36,8 @@ int main(int argc, char* argv[]) {
   APFEL::SetMaxFlavourPDFs(5);
 
   // Next we read the **APPLgrid** through the usual procedure
-  appl::grid g("./tests/atlas-Z0-rapidity.root");
+  const std::string gridfile = "./tests/atlas-Z0-rapidity.root";
+  const appl::grid  g(gridfile);
 
   // Once the **APPLgrid** is read, we can specify the initial-scale grid of x-points
   // to be used for the interpolation of the evolution factors. To do this
@@ -51,21 +52,17 @@ int main(int argc, char* argv[]) {
   // scale of *Q0 = 1 GeV*.
   const double Q0 = 1.0;
 
-  // Next we must specify if the **APPLgrid** has the pdf weighting option enabled.
-  // It is not currently possible to judge this from **APFELgrid**, so it must be specified externally.
-  const bool pdfweight = false;
-
   // The **FK** table is then generated with a single call to computeFK. Its arguments are
+  // + *Q0*           - the requested initial scale
   // + *"ATLASZRAP"*  - the requested name of the **FK** table
   // + *g*            - the **APPLgrid** to be combined
-  // + *Q0*           - the requested initial scale
-  // + *pdfweight*    - the boolean specifying whether or not the grid is weighted.
+  // + *gridfile*     - the path to the **APPLgrid** to be combined (required to reconstruct some APPLgrid parameters)
 
   // *computeFK* will then return an *NNPDF::FKTable<double>*. The template parameter specifies
   // the precision of the internal representation of the FK table, along with the precision of its
   // output after convolution. For generation purposes this is fixed at double precision, however
-  // the grid may be later read at single precision to enjoy the benefits of faster convolution.
-  NNPDF::FKTable<double>* FK = APFELgrid::computeFK("ATLASZRAP", g, Q0, pdfweight);
+  // the grid may be later read at single precision in order to enjoy the benefits of faster convolution.
+  NNPDF::FKTable<double>* FK = APFELgrid::computeFK(Q0, "ATLASZRAP", g, gridfile);
 
   // Now the user may add various tags to the FK table as they wish, to help identify e.g
   // the theory parameters used in its generation later. For example:

### 2016-04-20 tests/example_gen.cc
notation

@@ -48,17 +48,17 @@ int main(int argc, char* argv[]) {
   APFEL::SetGridParameters(2,15,5,1e-1);
 
   // With the x-grid specified, we can now generate the **FK** table. Firstly we specify
-  // and initial scale from which the evolution is to be performed. Here we use an initial
+  // an initial scale from which the evolution is to be performed. Here we are using an initial
   // scale of *Q0 = 1 GeV*.
   const double Q0 = 1.0;
 
-  // The **FK** table is then generated with a single call to computeFK. Its arguments are
+  // The **FK** table is then generated with a single call to *computeFK*. Its arguments are
   // + *Q0*           - the requested initial scale
   // + *"ATLASZRAP"*  - the requested name of the **FK** table
   // + *g*            - the **APPLgrid** to be combined
-  // + *gridfile*     - the path to the **APPLgrid** to be combined (required to reconstruct some APPLgrid parameters)
+  // + *gridfile*     - the path to the **APPLgrid** to be combined
 
-  // *computeFK* will then return an *NNPDF::FKTable<double>*. The template parameter specifies
+  // *computeFK* will then return an *NNPDF::FKTable\<double\>*. The template parameter specifies
   // the precision of the internal representation of the FK table, along with the precision of its
   // output after convolution. For generation purposes this is fixed at double precision, however
   // the grid may be later read at single precision in order to enjoy the benefits of faster convolution.

### 2016-04-20 paper.tex
editing the paper

@@
   posteriori} by means of interpolation techniques. Examples of the
-implementation of such a strategy are {\tt
-  APPLgrid}~\cite{Carli:2010rw} and {\tt
-  FastNLO}~\cite{Wobisch:2011ij}.
+implementation of this strategy are {\tt APPLgrid}~\cite{Carli:2010rw}
+and {\tt FastNLO}~\cite{Wobisch:2011ij}. For the computation of the
+hard cross sections, these packages rely on external codes to which
+they are interfaced, more or less directly, by means of a suite of
+functions whose scope is that of filling up PDF- and
+$\alpha_s$-independent lookup tables of cross-section weights that can
+subsequently be combined with an arbitrary PDF set. Examples of codes that have been directly interfaced to
 

### 2016-04-20 paper.tex
Equation dump

@@
 % \biboptions{}
+
+\newcommand{\be}{\begin{equation}}
+\newcommand{\ee}{\end{equation}}
+\newcommand{\ba}{\begin{eqnarray}}
+\newcommand{\ea}{\end{eqnarray}}
 
@@
 
+\section{The FastKernel method}
+
+\be  \sigma_{pp\to X} = \left( \frac{\alpha_s(Q^2)}{2\pi} \right)^{p} \int dx_1\, dx_2\,  f_i(x_1,Q^2)\; d\hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2) \, . \label{eq:hadconv} \ee
+\be  \sigma_{pp\to X} = \left( \frac{\alpha_s(Q^2)}{2\pi} \right)^{p} \sum_{\alpha,\beta}^{N_X}  f_i(x_\alpha,Q^2)\; W_{\alpha\beta,ij}\; f_j(x_\beta,Q^2) \, . \label{eq:hadconv2} \ee
+\be W_{\alpha\beta,ij} =  \int dx_1\, dx_2\,  \mathcal{I}^{(\alpha)}(x_1)\; d\hat{\sigma}_{ij\to X}\; \mathcal{I}^{(\beta)}(x_2) \, . \ee
+\be
+\label{eq:applconv}
+\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
+F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right),
+\ee
+\be F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) =  \sum_{i,j}^{13} C^{(s)}_{ij}  \left( f_i(x_{\alpha},Q^2_\tau)f_j(x_{\beta},Q^2_\tau) \right).  \label{eq:APPLsubproc}\ee
+\be
+f_i(x_{\alpha},Q^2_\tau) = \sum_{k}^{n_f} \sum_\beta^{N_x} A^\tau_{\alpha\beta ik}\; f_k(x_\beta,Q^2_0). \label{eq:fastPDFfinal_recalled}
+\ee 
+\ba F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) &=&  \sum_{i,j}^{13} \sum_{k,l}^{n_f}  \sum_{\delta,\gamma}^{N_x} C^{(s)}_{ij}  \left[  A^\tau_{\alpha\delta ik}\; f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\; \\
+&=&   \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0), \label{eq:FK1}
+\ea
+\be \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} = \sum_{i,j}^{13} C^{(s)}_{ij} A^\tau_{\alpha\delta ik} A^\tau_{\beta\gamma jl}.\ee
+\be
+\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
+ \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0).
+\ee
+\be
+\label{eq:FK}
+\sigma = \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} 
+\widetilde{W}_{kl\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
+\ee
+\be
+\label{eq:FKTable}
+\widetilde{W}_{kl\delta\gamma} = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} .
+\ee
+
 %% The Appendices part is started with the command \appendix;

### 2016-04-20 paper.tex
first draft of procedure section

--- 2016-04-20 paper.tex
@@
 
+\section{The FastKernel method}
+Hadron collider observables are typically computed in QCD by a double convolution of parton densities with a hard scattering cross-section.
+Consider for example the calculation of a general cross section $pp\to X$ with a set of parton distribution functions $\{f\}$ up to order $N_\alpha$ in the strong coupling
+\be  \sigma_{pp\to X} = \sum_{p}^{N_\alpha}\left( \frac{\alpha_s(Q^2)}{2\pi} \right)^{p+p_{\mathrm{LO}}} \int dx_1\, dx_2\,  f_i(x_1,Q^2)\; d\hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2) \, \label{eq:hadconv} \ee
+where $p_{\mathrm{LO}}$ is the leading order power of the strong coupling for the process $pp\to X$, $Q^2$ is the hard scale of the process and the indices $i,j$ sum over the active partonic species in the calculation.
+The central observation of tools such as {\tt APPLgrid} and {\tt fastNLO} is that such a calculation can be expressed as a more computationally efficient product
+\be
+\label{eq:applconv}
+\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
+F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right),
+\ee
+in terms of a precomputed set of coefficients $W$ obtained by convolving the hard scattering cross-section with a set of interpolating functions in the scale of the process and the parton-$x$ of each incoming parton.
+In Eqn.
+\ref{eq:applconv} the indices $\alpha, \beta$ run over points in the $x$-space interpolation grid, $\tau$ runs over the factorisation/renormalisation scale and $s$ runs over the (process-specific) parton subprocess densities defined as
+\be F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) =  \sum_{i,j}^{13} C^{(s)}_{ij}  \left( f_i(x_{\alpha},Q^2_\tau)f_j(x_{\beta},Q^2_\tau) \right).  \label{eq:APPLsubproc}\ee
+Additionally a number of tools (e.g {\tt APFEL}, {\tt HOPPET}) are available which perform PDF evolution via an analogous procedure.
+PDFs at some scale $Q^2_\tau$ evaluated upon an interpolation grid can be calculated from a set of PDFs at some initial scale $Q^2_0$ as
+\be
+f_i(x_{\alpha},Q^2_\tau) = \sum_{k}^{n_f} \sum_\beta^{N_x} A^\tau_{\alpha\beta ik}\; f_k(x_\beta,Q^2_0). \label{eq:fastPDFfinal_recalled}
+\ee 
+where the table $A$ interpolates DGLAP evolution kernels upon an interpolation grid in $x$.
+Given these interpolated coefficients, we may replace the (general-scale) PDFs used in the subprocess parton density Eqn.
+\ref{eq:APPLsubproc} with their equivalent expressions c.f Eqn.
+\ref{eq:fastPDFfinal_recalled} like so
+\ba F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) &=&  \sum_{i,j}^{13} \sum_{k,l}^{n_f}  \sum_{\delta,\gamma}^{N_x} C^{(s)}_{ij}  \left[  A^\tau_{\alpha\delta ik}\; f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\; \\
+&=&   \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0), \label{eq:FK1}
+\ea
+where the object
+\be \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} = \sum_{i,j}^{13} C^{(s)}_{ij} A^\tau_{\alpha\delta ik} A^\tau_{\beta\gamma jl}.\ee
+combines the operations of subprocess density construction and PDF evolution.
+Going further it is possible to then use the expression for subprocess parton densities in Eqn.~\ref{eq:FK1} in the full cross-section calculation:
+\be
+\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
+ \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0).
+\ee
+Performing some further contractions it is possible to obtain an extremely compact expression for the calculation of the cross-section in question, in terms of only the initial-scale PDFs and summing only over the initial scale interpolating $x$-grid and incoming parton flavors.
+\be
+\label{eq:FK}
+\sigma = \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} 
+\widetilde{W}_{kl\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
+\ee
+where the object
+\be
+\label{eq:FKTable}
+\widetilde{W}_{kl\delta\gamma} = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
+W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} .
+\ee
+is referred to here as an {\tt FK} table, and combines the information stored in {\tt APPLgrid}-style interpolated weight grids with analogously interpolated DGLAP evolution kernels.
+Such a combination
+enables for a maximally computationally-efficient expression for the calculation of observables at hadron colliders.
+
 %% The Appendices part is started with the command \appendix;

### 2016-04-20 paper.tex
editing the introduction

@@
 required. The main novelties are the precomputation of the PDF and
-$\alpha_s$ evolution and the optimization of the numerical convolution
-between hard cross sections and PDFs. We demostrate that our techique
+$\alpha_s$ evolution and the optimisation of the numerical convolution
+between hard cross sections and PDFs. We demonstrate that our technique
 can lead to a substantial speed improvement as compared to the
@@
 convolution with any set of PDFs can be quickly performed {\it a
-  posteriori} by means of interpolation techniques. Examples of the
-implementation of this strategy are {\tt APPLgrid}~\cite{Carli:2010rw}
-and {\tt FastNLO}~\cite{Wobisch:2011ij}. For the computation of the
-hard cross sections, these packages rely on external codes to which
-they are interfaced, more or less directly, by means of a suite of
+  posteriori} by means of interpolation techniques.
+
+Examples of the implementation of this strategy are {\tt
+  APPLgrid}~\cite{Carli:2010rw} and {\tt
+  FastNLO}~\cite{Wobisch:2011ij}. For the computation of the hard
+cross sections, these packages rely on external codes to which they
+are interfaced, more or less directly, by means of a suite of
 functions whose scope is that of filling up PDF- and
-$\alpha_s$-independent lookup tables of cross-section weights that can
-subsequently be combined with an arbitrary PDF set. Examples of codes that have been directly interfaced to
+$\alpha_s$-independent look-up tables of cross-section weights that can
+subsequently be combined with an arbitrary set of PDFs. Examples of
+codes that have been directly interfaced to {\tt APPLgrid} and/or to
+{\tt FastNLO} are {\tt MCFM}~\cite{Campbell:2010ff} and {\tt
+  NLOJet++}~\cite{Nagy:2003tz}. More recently, dedicated interfaces to
+automated general-purpose Monte Carlo (MC) event generators have been
+developed. In particular, the {\tt aMCfast} and {\tt MCgrid} codes
+have been specifically designed to extract the relevant information
+from the {\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
+  SHERPA}~\cite{Gleisberg:2008ta} event generators, respectively, and
+fill up look-up interpolation tables in the {\tt APPLgrid} and/or {\tt
+  FastNLO} format.
 
-\section{The FastKernel method}
-Hadron collider observables are typically computed in QCD by a double convolution of parton densities with a hard scattering cross-section.
-Consider for example the calculation of a general cross section $pp\to X$ with a set of parton distribution functions $\{f\}$ up to order $N_\alpha$ in the strong coupling
-\be  \sigma_{pp\to X} = \sum_{p}^{N_\alpha}\left( \frac{\alpha_s(Q^2)}{2\pi} \right)^{p+p_{\mathrm{LO}}} \int dx_1\, dx_2\,  f_i(x_1,Q^2)\; d\hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2) \, \label{eq:hadconv} \ee
-where $p_{\mathrm{LO}}$ is the leading order power of the strong coupling for the process $pp\to X$, $Q^2$ is the hard scale of the process and the indices $i,j$ sum over the active partonic species in the calculation.
-The central observation of tools such as {\tt APPLgrid} and {\tt fastNLO} is that such a calculation can be expressed as a more computationally efficient product
-\be
+While these tools have proven to be extremely useful for the
+extraction of parton densities, the great amount of experimental data
+that has become available thanks to the LHC and that are or are going
+to be included into the current PDF analyses is stretching the
+capabilities of the general strategy as well as its
+implementations. In fact, presently a typical global PDF fit includes
+thousands of hadronic data points whose predictions have to be
+computed thousands of times during the minimisation process. As a
+consequence, the ``standard'' tools, $i.e.$ {\tt APPLgrid} and {\tt
+  FastNLO}, might not be fast enough to meet the needs of the modern
+PDF fitters. For this reason we believe that a high-performance tool
+tailored for PDF analyses is becoming increasingly desirable.
+
+The so-called {\tt FastKernel} method, developed in the framework of
+the NNPDF global analyses~\cite{Ball:2010de}, was conceived to address
+this problem. The basic principle of this method is the maximisation
+of the precomputation with the consequent minimisation of the amount
+of operations needed to be performed during the PDF fit. In
+particular, the {\tt FastKernel} method relies on the precomputation
+of the partonic hard cross sections and the DGLAP evolution kernels
+and their combination into a single look-up table, which we call {\tt
+  FKtable}, in such a way that the prediction for a given hadronic
+observable can be obtained by a single numerical convolution between
+the respective {\tt FKtable} and the initial scale PDFs, which are
+the quantities that are usually fitted to data.
+
+In this paper we present a public implementation of the {\tt
+  FastKernel} method in which the partonic hard cross sections are
+taken from any given {\tt APPLgrid} look-up table while the DGLAP
+evolution kernels are provided by {\tt APFEL}
+code~\cite{Bertone:2013vaa}. We named the resulting code {\tt
+  APFELgrid}.
+
+The outline of this paper is the following. In
+Sect.~\ref{sec:FastKernel} we present the tecnical details of the
+implemantation of the {\tt FastKernel} method.
+
+\section{The FastKernel method}\label{sec:FastKernel}
+
+Hadron collider observables are typically computed in QCD by a double
+convolution of parton densities with a hard scattering
+cross-section. Consider for example the calculation of a general cross
+section $pp\to X$ with a set of parton distribution functions $\{f\}$
+up to order $N_\alpha$ in the strong coupling \be \sigma_{pp\to X} =
+\sum_{p}^{N_\alpha}\left( \frac{\alpha_s(Q^2)}{2\pi}
+\right)^{p+p_{\mathrm{LO}}} \int dx_1\, dx_2\, f_i(x_1,Q^2)\;
+d\hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2) \, \label{eq:hadconv} \ee where
+$p_{\mathrm{LO}}$ is the leading order power of the strong coupling
+for the process $pp\to X$, $Q^2$ is the hard scale of the process and
+the indices $i,j$ sum over the active partonic species in the
+calculation. The central observation of tools such as {\tt APPLgrid}
+and {\tt FastNLO} is that such a calculation can be expressed as a
+more computationally efficient product \be
 \label{eq:applconv}

### 2016-04-20 paper.tex
editing the introduction

@@
 taken from any given {\tt APPLgrid} look-up table while the DGLAP
-evolution kernels are provided by {\tt APFEL}
+evolution kernels are provided by the {\tt APFEL}
 code~\cite{Bertone:2013vaa}. We named the resulting code {\tt

### 2016-04-22 paper.tex
adding some plot layout
updating plot
editing FK section

@@
 
+\clearpage
+
+\tableofcontents
+
 %% main text
@@
 Hadron collider observables are typically computed in QCD by a double
-convolution of parton densities with a hard scattering
-cross-section. Consider for example the calculation of a general cross
-section $pp\to X$ with a set of parton distribution functions $\{f\}$
-up to order $N_\alpha$ in the strong coupling \be \sigma_{pp\to X} =
-\sum_{p}^{N_\alpha}\left( \frac{\alpha_s(Q^2)}{2\pi}
-\right)^{p+p_{\mathrm{LO}}} \int dx_1\, dx_2\, f_i(x_1,Q^2)\;
-d\hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2) \, \label{eq:hadconv} \ee where
-$p_{\mathrm{LO}}$ is the leading order power of the strong coupling
-for the process $pp\to X$, $Q^2$ is the hard scale of the process and
-the indices $i,j$ sum over the active partonic species in the
-calculation. The central observation of tools such as {\tt APPLgrid}
-and {\tt FastNLO} is that such a calculation can be expressed as a
-more computationally efficient product \be
-\label{eq:applconv}
-\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x} \sum_{\tau}^{N_{Q}}
-W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
-F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right),
-\ee
-in terms of a precomputed set of coefficients $W$ obtained by convolving the hard scattering cross-section with a set of interpolating functions in the scale of the process and the parton-$x$ of each incoming parton.
-In Eqn.
-\ref{eq:applconv} the indices $\alpha, \beta$ run over points in the $x$-space interpolation grid, $\tau$ runs over the factorisation/renormalisation scale and $s$ runs over the (process-specific) parton subprocess densities defined as
-\be F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) =  \sum_{i,j}^{13} C^{(s)}_{ij}  \left( f_i(x_{\alpha},Q^2_\tau)f_j(x_{\beta},Q^2_\tau) \right).  \label{eq:APPLsubproc}\ee
-Additionally a number of tools (e.g {\tt APFEL}, {\tt HOPPET}) are available which perform PDF evolution via an analogous procedure.
-PDFs at some scale $Q^2_\tau$ evaluated upon an interpolation grid can be calculated from a set of PDFs at some initial scale $Q^2_0$ as
-\be
-f_i(x_{\alpha},Q^2_\tau) = \sum_{k}^{n_f} \sum_\beta^{N_x} A^\tau_{\alpha\beta ik}\; f_k(x_\beta,Q^2_0). \label{eq:fastPDFfinal_recalled}
-\ee 
-where the table $A$ interpolates DGLAP evolution kernels upon an interpolation grid in $x$.
-Given these interpolated coefficients, we may replace the (general-scale) PDFs used in the subprocess parton density Eqn.
-\ref{eq:APPLsubproc} with their equivalent expressions c.f Eqn.
-\ref{eq:fastPDFfinal_recalled} like so
-\ba F^{(s)}\left(x_{\alpha}, x_{\beta},  Q^2_{\tau}\right) &=&  \sum_{i,j}^{13} \sum_{k,l}^{n_f}  \sum_{\delta,\gamma}^{N_x} C^{(s)}_{ij}  \left[  A^\tau_{\alpha\delta ik}\; f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\; \\
-&=&   \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0), \label{eq:FK1}
-\ea
-where the object
-\be \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} = \sum_{i,j}^{13} C^{(s)}_{ij} A^\tau_{\alpha\delta ik} A^\tau_{\beta\gamma jl}.\ee
-combines the operations of subprocess density construction and PDF evolution.
-Going further it is possible to then use the expression for subprocess parton densities in Eqn.~\ref{eq:FK1} in the full cross-section calculation:
-\be
-\sigma = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
-W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}
- \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0).
-\ee
-Performing some further contractions it is possible to obtain an extremely compact expression for the calculation of the cross-section in question, in terms of only the initial-scale PDFs and summing only over the initial scale interpolating $x$-grid and incoming parton flavors.
-\be
-\label{eq:FK}
-\sigma = \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x} 
-\widetilde{W}_{kl\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
-\ee
-where the object
-\be
-\label{eq:FKTable}
-\widetilde{W}_{kl\delta\gamma} = \sum_p \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}^{N_x^\prime} \sum_{\tau}^{N_{Q}}
-W_{\alpha\beta\tau}^{(p)(s)} \, \left( \frac{\alpha_s\left(Q^2_{\tau}\right)}{2\pi}\right)^{p}\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} .
-\ee
-is referred to here as an {\tt FK} table, and combines the information stored in {\tt APPLgrid}-style interpolated weight grids with analogously interpolated DGLAP evolution kernels.
-Such a combination
-enables for a maximally computationally-efficient expression for the calculation of observables at hadron colliders.
+convolution of parton densities with a hard-scattering cross
+section. Consider for example the calculation of a general cross
+section $pp\to X$ with a set of PDFs $\{f\}$:
+\begin{equation} \label{eq:hadconv}
+  \sigma_{pp\to X} =
+  \sum_{i,j}\int dx_1\, dx_2\, f_i(x_1,Q^2)\;
+  \hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2)\,,
+\end{equation}
+where $Q^2$ is the typical hard scale of the process and the indices
+$i,j$ sum over the active partonic species in the calculation. The
+partonic cross section $\hat{\sigma}_{ij\to X}$ admits the
+perturbative expansion:
+\begin{equation}\label{eq:expansion}
+\hat{\sigma}_{ij\to X} = \alpha_s^{p_{\rm LO}}(Q^2) \sum_{p=0}^{N_\alpha} \alpha_s^{p}(Q^2)\hat{\sigma}^{(p)}_{ij\to X}
+\end{equation}
+$p_{\rm LO}$ being the $\alpha_s$ power of the leading-order
+contribution and $\hat{\sigma}^{(p)}_{ij\to X}$ the N$^p$LO
+contribution. Plugging Eq.~(\ref{eq:expansion}) into
+Eq.~(\ref{eq:hadconv}), one gets:
+\begin{equation}\label{eq:pertexp}
+  \sigma_{pp\to X} =
+  \sum_{i,j}\sum_{p=0}^{N_\alpha} \int dx_1\,dx_2\,
+  \hat{\sigma}^{(p)}_{ij\to X}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)\,.
+\end{equation}
+
+The central observation of tools such as {\tt APPLgrid} and {\tt
+  FastNLO} is that the part of the calculation depending on $\alpha_s$
+and PDFs can be computed by means of interpolation techniquies
+starting from the knowledge of these quantities for a finite number of
+values of $x_1$, $x_2$, and $Q^2$. In particular, one can use the
+interpolation formula:
+\begin{equation}\label{eq:interpolation}
+\begin{array}{c}
+\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)
+  =\\
+\\
+\displaystyle \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
+  \,f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)
+  \,\mathcal{I}_\tau^{(n')}(Q^2)\,\mathcal{I}_\alpha^{(n_1)}(x_1)
+  \,\mathcal{I}_\beta^{(n_2)}(x_2)
+\end{array}
+\end{equation}
+where $\mathcal{I}_\tau^{(n')}(Q^2)$,
+$\mathcal{I}_\alpha^{(n_1)}(x_1)$, $\mathcal{I}_\beta^{(n_2)}(x_2)$
+are interpolating functions of degree $n'$, $n_1$, and $n_2$
+(typically Lagrange polynomials) over three independent interpolation
+grids in $Q^2$, $x_1$, and $x_2$. Plugging
+Eq.~(\ref{eq:interpolation}) into Eq.~(\ref{eq:pertexp}), one finds:
+\begin{equation}\label{eq:XsecPDFs}
+\begin{array}{rcl}
+  \displaystyle \sigma_{pp\to X} &=& \displaystyle 
+                                     \sum_{i,j}\sum_{p=0}^{N_\alpha} \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
+                                     \,f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)\\
+  \\
+                                 &\times& \displaystyle \mathcal{I}_\tau^{(n')}(Q^2)\int dx_1\,dx_2\,
+                                          \hat{\sigma}^{(p)}_{ij\to X}\,\mathcal{I}_\alpha^{(n_1)}(x_1)
+                                          \,\mathcal{I}_\beta^{(n_2)}(x_2)\,.
+\end{array}
+\end{equation}
+
+In addition, usually not all possible pairs $(i,j)$ in
+Eq.~(\ref{eq:hadconv}) contribute to the hadronic cross section and
+only a limited number of partonic subprocesses and thus one do the
+following replacement:
+\begin{equation}
+\sum_{i,j} \hat{\sigma}^{(p)}_{ij\to X}
+f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)\rightarrow
+\sum_{s}^{N_{\rm sub}} \hat{\sigma}^{(p)(s)} F^{(s)}_{\alpha\beta,\tau}
+\end{equation}
+where the $s$ index runs over the $N_{\rm sub}$ (process-specific)
+parton subprocess:
+\begin{equation}\label{eq:APPLsubproc}
+  F^{(s)}_{\alpha\beta,\tau} =\sum_{i,j} C^{(s)}_{ij} 
+  f_i(x_{\alpha},Q^2_\tau)f_j(x_{\beta},Q^2_\tau)\,,
+\end{equation}
+where the $C^{(s)}_{ij}$ matrix encodes the combination of PDFs that
+contribute to the $s$-th subprocess and $\hat{\sigma}^{(p)(s)}$
+is the corresponding hard cross section. Finally, defining:
+\begin{equation}
+  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau^{(n')}(Q^2)\int dx_1\,dx_2\,
+  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha^{(n_1)}(x_1)
+  \,\mathcal{I}_\beta^{(n_2)}(x_2)\,,
+\end{equation}
+one can write Eq.~(\ref{eq:XsecPDFs}) as follows:
+\begin{equation} \label{eq:applconv}
+  \sigma_{pp\to X} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta,\tau} 
+  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
+\end{equation}
+
+Additionally, a number of tools, $e.g.$ {\tt
+  APFEL}~\cite{Bertone:2013vaa} or {\tt HOPPET}~\cite{Salam:2008qg},
+are available which perform PDF evolution via an analogous
+procedure. {\tt APFEL} in particular, is able to provide the so-called
+\textit{evolution operator}. The evolution operator, is computed
+assuming that the DGLAP evolution can be encoded in a single operator
+whose action on the intial scale PDFs proceduces the final scale
+PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
+on the $x$-space grid one can write:
+\begin{equation}\label{eq:fastPDFfinal_recalled}
+  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
+  \sum_\beta A^\tau_{\alpha\beta, ik}\;
+  f_k(x_\beta,Q^2_0)\,, 
+\end{equation}
+where the operator $A$ is the solution of the DGLAP evolution
+equation with baundary condition:
+\begin{equation}
+A^\tau_{\alpha\beta, ik} =
+\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
+\end{equation}
+
+Given the evolution operator, we may replace the (general-scale) PDFs
+used in the subprocess parton density Eq.~(\ref{eq:APPLsubproc}) with
+their equivalent expressions c.f Eq.~(\ref{eq:fastPDFfinal_recalled})
+like so:
+\begin{equation}\label{eq:FK1}
+\begin{array}{rcl}
+F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
+                               \sum_{\delta,\gamma} C^{(s)}_{ij}
+                               \left[  A^\tau_{\alpha\delta ik}\;
+                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
+                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
+  \\
+\\
+&=& \displaystyle \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x}
+\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
+f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
+\end{array}
+\end{equation}
+where the object:
+\begin{equation}
+  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
+  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
+  A^\tau_{\beta\gamma jl}
+\end{equation}
+combines the operations of subprocess density construction and PDF
+evolution. Going further it is possible to then use the expression for
+subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
+cross-section calculation:
+\begin{equation}
+\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p^{N_\alpha}
+\sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}
+\sum_{\tau} 
+\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
+f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
+\end{equation}
+Performing some further contractions it is possible to obtain an
+extremely compact expression for the calculation of the cross section
+in question, in terms of only the initial-scale PDFs and summing only
+over the initial scale interpolating $x$-grid and the incoming parton
+flavors:
+\begin{equation}\label{eq:FK}
+  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
+  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
+\end{equation}
+where the object:
+\begin{equation}\label{eq:FKTable}
+  \widetilde{W}_{kl,\delta\gamma} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta} \sum_{\tau}
+\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
+\end{equation}
+is referred to here as an {\tt FKtable}, and combines the information
+stored in {\tt APPLgrid}-style interpolated weight grids with
+analogously interpolated DGLAP evolution operators. Such a combination
+enables for a maximally computationally-efficient expression for the
+calculation of observables at hadron colliders.
+
+\section{{\tt APFELgrid} manual}
+
+\section{Benchmark}
+
+In this section we discuss...
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[ph]
+\centering
+\includegraphics[scale=0.5]{plots/t0}
+\caption{\small Performance comparisons between {\tt APFELgrid} with
+  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
+  convolution time per point and process.}
+\label{fig:benchmark}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-04-28 paper.tex
first pass on introduction

@@
 \begin{abstract}
-We present a new method conceived to ease the inclusion of hadronic
-observables into parton distribution function (PDF) fits. Our
-technique relies on the the same principles adopted by the existing
-fast interfaces but implements a set of improvements aimed at
-optimizing the performance in the context of PDF fits where many
-iterations and many predictions for each iteration are usually
-required. The main novelties are the precomputation of the PDF and
-$\alpha_s$ evolution and the optimisation of the numerical convolution
-between hard cross sections and PDFs. We demonstrate that our technique
+We present a new software package designed to aid the inclusion of hadronic
+observables into Parton Distribution Function (PDF) fits. The {\tt APFELgrid} package
+converts interpolated weight tables provided by {\tt APPLgrid} files into a more efficient
+format for PDF fitting by the pre-computation and combination with PDF and $\alpha_s$ evolution factors provided
+by {\tt APFEL}. This combination significantly reduces the number of operations required to perform the calculation
+of hadronic observables in PDF fits and simplifies the structure of the calculation into
+a readily optimised scalar product. We demonstrate that our technique
 can lead to a substantial speed improvement as compared to the
-existing interfaces.
+existing interfaces without any reduction in numerical accuracy.
+
+%We present a new method conceived to ease the inclusion of hadronic
+%observables into parton distribution function (PDF) fits. Our
+%technique relies on the the same principles adopted by the existing
+%fast interfaces but implements a set of improvements aimed at
+%optimizing the performance in the context of PDF fits where many
+%iterations and many predictions for each iteration are usually
+%required. The main novelties are the precomputation of the PDF and
+%$\alpha_s$ evolution and the optimisation of the numerical convolution
+%between hard cross sections and PDFs. We demonstrate that our technique
+%can lead to a substantial speed improvement as compared to the
+%existing interfaces.
 \end{abstract}
@@
   %Leave blank, supplied by Elsevier.
-{\em Licensing provisions:}                                   \\
-  %enter "none" if CPC non-profit use license is sufficient.
-{\em Programming language:}                                   \\
-{\em Computer:}                                               \\
+{\em Licensing provisions:}   MIT license                                \\
+{\em Programming language:}  C++                                 \\
+{\em Computer:} PC/Mac                                               \\
   %Computer(s) for which program has been designed.
-{\em Operating system:}                                       \\
+{\em Operating system:} MacOS/Linux                                       \\
   %Operating system(s) for which program has been designed.
-{\em RAM:} bytes                                              \\
+{\em RAM:} varying                                              \\
   %RAM in bytes required to execute program with typical data.
-{\em Number of processors used:}                              \\
-  %If more than one processor.
-{\em Supplementary material:}                                 \\
-  % Fill in if necessary, otherwise leave out.
-{\em Keywords:} Keyword one, Keyword two, Keyword three, etc.  \\
+{\em Keywords:} QCD, PDF\\
   % Please give some freely chosen keywords that we can use in a
   % cumulative keyword index.
-{\em Classification:}                                         \\
+{\em Classification:}  11.1 General, High Energy Physics and Computing                                       \\
   %Classify using CPC Program Library Subject Index, see (
@@
   %e.g. 4.4 Feynman diagrams, 5 Computer Algebra.
-{\em External routines/libraries:}                            \\
+{\em External routines/libraries:}  {\tt APPLgrid}, {\tt APFEL}                          \\
   % Fill in if necessary, otherwise leave out.
-{\em Subprograms used:}                                       \\
-  %Fill in if necessary, otherwise leave out.
-{\em Catalogue identifier of previous version:}*              \\
-  %Only required for a New Version summary, otherwise leave out.
-{\em Journal reference of previous version:}*                  \\
-  %Only required for a New Version summary, otherwise leave out.
-{\em Does the new version supersede the previous version?:}*   \\
-  %Only required for a New Version summary, otherwise leave out.
 {\em Nature of problem:}\\
-  %Describe the nature of the problem here.
+ Fast computation of hadronic observables under the variation of parton distribution functions.
    \\
 {\em Solution method:}\\
-  %Describe the method solution here.
+  Combination of interpolated weight grids from {\tt APPLgrid} and evolution factors from {\tt APFEL} into efficient {\tt FastKernel} tables.\\
+{\em Running time:} varying\\
    \\
-{\em Reasons for the new version:}*\\
-  %Only required for a New Version summary, otherwise leave out.
-   \\
-{\em Summary of revisions:}*\\
-  %Only required for a New Version summary, otherwise leave out.
-   \\
-{\em Restrictions:}\\
-  %Describe any restrictions on the complexity of the problem here.
-   \\
-{\em Unusual features:}\\
-  %Describe any unusual features of the program/problem here.
-   \\
-{\em Additional comments:}\\
-  %Provide any additional comments here.
-   \\
-{\em Running time:}\\
-  %Give an indication of the typical running time here.
-   \\
-* Items marked with an asterisk are only required for new versions
-of programs previously published in the CPC Program Library.\\
 \end{small}
@@
 
-\tableofcontents
+%%\tableofcontents
 
@@
 
-The need for fast and accurate cross-section predictions in the Large
-Hadron Collider (LHC) era is becoming increasingly important. In fact,
-the outstanding amount of precise data produced by the LHC experiments
-is bound to provide a very stringent constraint on the proton parton
-distribution functions (PDFs) on the condition that as many
-measurements as possible are included in PDF fits with the highest
-accuracy possible. However, the direct use of most of the codes that
-provide predictions for non-inclusive hadron-collider observables in a
-PDF fit, which typically requires a large number of iterations, is not
-viable because the time required to obtain accurate predictions is
-usually of the order of a few hours or more. In order to overcome this
-limitation, the typical strategy adopted relies on the precomputation
-of the partonic hard cross sections in such a way that the numerical
-convolution with any set of PDFs can be quickly performed {\it a
-  posteriori} by means of interpolation techniques.
+The ability to perform fast and accurate cross-section predictions for measurements at
+the Large Hadron Collider (LHC) is a vital requirement for the most precise PDF determinations. LHC measurements have a unique
+capacity to shed light upon the dynamics of the proton and provide stringent constraints upon 
+the proton PDF. In order to make full use of current and future LHC data, efficient strategies
+for the computation of predictions for these observables must be deployed. Software packages performing predictions 
+for non-inclusive hadron-collider observables cannot be easily deployed in a PDF fit requiring rapid iteration due the time required to obtain accurate results (usually of the order of a few hours or more
+per data point). In order to overcome this limitation, the typical strategy adopted for fast cross-section prediction relies on the precomputation
+of the partonic hard cross sections in such a way that the numerical convolution with any set of PDFs can be quickly performed by means of interpolation techniques.
 
-Examples of the implementation of this strategy are {\tt
+This interpolation strategy is implemented in the {\tt
   APPLgrid}~\cite{Carli:2010rw} and {\tt
-  FastNLO}~\cite{Wobisch:2011ij}. For the computation of the hard
+  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the hard
 cross sections, these packages rely on external codes to which they
-are interfaced, more or less directly, by means of a suite of
-functions whose scope is that of filling up PDF- and
-$\alpha_s$-independent look-up tables of cross-section weights that can
-subsequently be combined with an arbitrary set of PDFs. Examples of
-codes that have been directly interfaced to {\tt APPLgrid} and/or to
-{\tt FastNLO} are {\tt MCFM}~\cite{Campbell:2010ff} and {\tt
-  NLOJet++}~\cite{Nagy:2003tz}. More recently, dedicated interfaces to
+are interfaced by means of a suite of
+functions allowing for the filling of PDF- and
+$\alpha_s$-independent look-up tables of cross-section weights. Codes
+such as {\tt MCFM}~\cite{Campbell:2010ff} and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to 
+ {\tt APPLgrid}/{\tt FastNLO} and more recently dedicated interfaces to
 automated general-purpose Monte Carlo (MC) event generators have been
-developed. In particular, the {\tt aMCfast} and {\tt MCgrid} codes
+developed. The {\tt aMCfast} and {\tt MCgrid} codes
 have been specifically designed to extract the relevant information
 from the {\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
-  SHERPA}~\cite{Gleisberg:2008ta} event generators, respectively, and
-fill up look-up interpolation tables in the {\tt APPLgrid} and/or {\tt
-  FastNLO} format.
+  SHERPA}~\cite{Gleisberg:2008ta} event generators respectively.
 
-While these tools have proven to be extremely useful for the
-extraction of parton densities, the great amount of experimental data
-that has become available thanks to the LHC and that are or are going
-to be included into the current PDF analyses is stretching the
-capabilities of the general strategy as well as its
-implementations. In fact, presently a typical global PDF fit includes
-thousands of hadronic data points whose predictions have to be
+While these tools have proven to be extremely useful in the
+extraction of parton densities, the volume of experimental data
+made available by LHC collaborations for use in PDF fits is already
+stretching the capabilities of typical fitting technology. Presently a typical global PDF fit includes
+thousands of hadronic data points for which predictions have to be
 computed thousands of times during the minimisation process. As a
-consequence, the ``standard'' tools, $i.e.$ {\tt APPLgrid} and {\tt
-  FastNLO}, might not be fast enough to meet the needs of the modern
-PDF fitters. For this reason we believe that a high-performance tool
-tailored for PDF analyses is becoming increasingly desirable.
+consequence the ``standard'' tools, $i.e.$ {\tt APPLgrid} and {\tt
+  FastNLO}, might not be fast enough to meet the needs of modern
+PDF fitters. For this reason a high-performance tool tailored specifically 
+for PDF analyses is becoming increasingly important.
 
-The so-called {\tt FastKernel} method, developed in the framework of
-the NNPDF global analyses~\cite{Ball:2010de}, was conceived to address
-this problem. The basic principle of this method is the maximisation
-of the precomputation with the consequent minimisation of the amount
-of operations needed to be performed during the PDF fit. In
-particular, the {\tt FastKernel} method relies on the precomputation
-of the partonic hard cross sections and the DGLAP evolution kernels
-and their combination into a single look-up table, which we call {\tt
-  FKtable}, in such a way that the prediction for a given hadronic
-observable can be obtained by a single numerical convolution between
-the respective {\tt FKtable} and the initial scale PDFs, which are
-the quantities that are usually fitted to data.
+The {\tt FastKernel} method, developed in the framework of
+the NNPDF global analyses~\cite{Ball:2010de}, was developed to address
+this problem. This method differs from the standard procedure \`{a} la {\tt APPLgrid} or {\tt fastNLO} in that it
+maximises the amount of information that is precomputed prior to fitting such as to minimise the amount
+of operations to be performed during the fit. Specifically, the {\tt FastKernel} method relies on the precomputation
+of the partonic hard cross sections and DGLAP evolution kernels
+and their combination into a single look-up table, here called a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction for a given hadronic
+observable can be obtained by a simple matrix product between
+the respective {\tt FK} table and the initial scale PDFs.
 
-In this paper we present a public implementation of the {\tt
-  FastKernel} method in which the partonic hard cross sections are
-taken from any given {\tt APPLgrid} look-up table while the DGLAP
-evolution kernels are provided by the {\tt APFEL}
-code~\cite{Bertone:2013vaa}. We named the resulting code {\tt
-  APFELgrid}.
+In this paper we present the {\tt APFELgrid} package, a public implementation of the {\tt
+  FastKernel} method in which the hard partonic cross sections provided in an {\tt APPLgrid} look-up table are combined with the DGLAP
+evolution kernels provided by the {\tt APFEL}
+package~\cite{Bertone:2013vaa}.
 

### 2016-04-28 paper.tex
short pass on FK section

@@
   \sigma_{pp\to X} =
-  \sum_{i,j}\int dx_1\, dx_2\, f_i(x_1,Q^2)\;
+  \sum_{i,j}^{N_F}\int dx_1\, dx_2\, f_i(x_1,Q^2)\;
   \hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2)\,,
@@
 partonic cross section $\hat{\sigma}_{ij\to X}$ admits the
-perturbative expansion:
+perturbative expansion
 \begin{equation}\label{eq:expansion}
@@
 contribution and $\hat{\sigma}^{(p)}_{ij\to X}$ the N$^p$LO
-contribution. Plugging Eq.~(\ref{eq:expansion}) into
-Eq.~(\ref{eq:hadconv}), one gets:
+contribution. Therefore to order $N_\alpha$ in QCD
 \begin{equation}\label{eq:pertexp}
   \sigma_{pp\to X} =
-  \sum_{i,j}\sum_{p=0}^{N_\alpha} \int dx_1\,dx_2\,
+  \sum_{i,j}^{N_F}\sum_{p=0}^{N_\alpha} \int dx_1\,dx_2\,
   \hat{\sigma}^{(p)}_{ij\to X}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)\,.
@@
 The central observation of tools such as {\tt APPLgrid} and {\tt
-  FastNLO} is that the part of the calculation depending on $\alpha_s$
-and PDFs can be computed by means of interpolation techniquies
-starting from the knowledge of these quantities for a finite number of
-values of $x_1$, $x_2$, and $Q^2$. In particular, one can use the
-interpolation formula:
+  FastNLO} is that the PDF and $\alpha_S$ dependence may be factorised out of
+  the convolution via expansion over a set of interpolating functions in $Q^2$ and the
+  two values of parton-$x$. Explicitly, one may represent the PDFs and $\alpha_S$ as
 \begin{equation}\label{eq:interpolation}
@@
 \end{equation}
-where $\mathcal{I}_\tau^{(n')}(Q^2)$,
+in terms of interpolating functions $\mathcal{I}_\tau^{(n')}(Q^2)$,
 $\mathcal{I}_\alpha^{(n_1)}(x_1)$, $\mathcal{I}_\beta^{(n_2)}(x_2)$
-are interpolating functions of degree $n'$, $n_1$, and $n_2$
-(typically Lagrange polynomials) over three independent interpolation
-grids in $Q^2$, $x_1$, and $x_2$. Plugging
-Eq.~(\ref{eq:interpolation}) into Eq.~(\ref{eq:pertexp}), one finds:
+(typically Lagrange polynomials) of degree $n'$, $n_1$, and $n_2$.
+Using these expressions for the PDFs in the double-convolution of Eq.~(\ref{eq:pertexp}) one finds:
 \begin{equation}\label{eq:XsecPDFs}
@@
 
-In addition, usually not all possible pairs $(i,j)$ in
-Eq.~(\ref{eq:hadconv}) contribute to the hadronic cross section and
-only a limited number of partonic subprocesses and thus one do the
-following replacement:
+Additionally one may take advantages of symmetries present in the matrix element to reduce the sum over active flavours $(i,j)$ in
+Eq.~(\ref{eq:hadconv}) to a reduced number of partonic subprocesses with a replacement such as
 \begin{equation}

### 2016-04-29 paper.tex
hspace for figure
adding split plot
more edits

@@
 Hadron collider observables are typically computed in QCD by a double
-convolution of parton densities with a hard-scattering cross
-section. Consider for example the calculation of a general cross
-section $pp\to X$ with a set of PDFs $\{f\}$:
-\begin{equation} \label{eq:hadconv}
-  \sigma_{pp\to X} =
-  \sum_{i,j}^{N_F}\int dx_1\, dx_2\, f_i(x_1,Q^2)\;
-  \hat{\sigma}_{ij\to X}\; f_j(x_2,Q^2)\,,
-\end{equation}
-where $Q^2$ is the typical hard scale of the process and the indices
-$i,j$ sum over the active partonic species in the calculation. The
-partonic cross section $\hat{\sigma}_{ij\to X}$ admits the
-perturbative expansion
-\begin{equation}\label{eq:expansion}
-\hat{\sigma}_{ij\to X} = \alpha_s^{p_{\rm LO}}(Q^2) \sum_{p=0}^{N_\alpha} \alpha_s^{p}(Q^2)\hat{\sigma}^{(p)}_{ij\to X}
-\end{equation}
-$p_{\rm LO}$ being the $\alpha_s$ power of the leading-order
-contribution and $\hat{\sigma}^{(p)}_{ij\to X}$ the N$^p$LO
-contribution. Therefore to order $N_\alpha$ in QCD
+convolution of parton densities with a hard scattering cross-section. Consider for example the calculation of a general cross-section $pp\to X$ with a set of PDFs $\{f\}$:
 \begin{equation}\label{eq:pertexp}
   \sigma_{pp\to X} =
-  \sum_{i,j}^{N_F}\sum_{p=0}^{N_\alpha} \int dx_1\,dx_2\,
+  \sum_{i,j}\sum_{p=0} \int dx_1\,dx_2\,
   \hat{\sigma}^{(p)}_{ij\to X}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)\,.
 \end{equation}
+where $Q^2$ is the typical hard scale of the process, the indices
+$i,j$ sum over the active partonic species in the calculation, $p_{\rm LO}$ is the leading-order power of $\alpha_s$ for the process
+and $\hat{\sigma}^{(p)}_{ij\to X}$ the N$^p$LO contribution.
 
@@
   FastNLO} is that the PDF and $\alpha_S$ dependence may be factorised out of
-  the convolution via expansion over a set of interpolating functions in $Q^2$ and the
-  two values of parton-$x$. Explicitly, one may represent the PDFs and $\alpha_S$ as
+  the convolution via expansion over a set of interpolating functions, spanning $Q^2$ and the
+  two values of parton-$x$. For example one may represent the PDFs and $\alpha_S$ as
 \begin{equation}\label{eq:interpolation}
@@
   \,f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)
-  \,\mathcal{I}_\tau^{(n')}(Q^2)\,\mathcal{I}_\alpha^{(n_1)}(x_1)
-  \,\mathcal{I}_\beta^{(n_2)}(x_2)
+  \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
+  \,\mathcal{I}_\beta(x_2)
 \end{array}
 \end{equation}
-in terms of interpolating functions $\mathcal{I}_\tau^{(n')}(Q^2)$,
-$\mathcal{I}_\alpha^{(n_1)}(x_1)$, $\mathcal{I}_\beta^{(n_2)}(x_2)$
-(typically Lagrange polynomials) of degree $n'$, $n_1$, and $n_2$.
+in terms of the Lagrange basis polynomials $\mathcal{I}_\tau(Q^2)$,
+$\mathcal{I}_\alpha(x_1)$, $\mathcal{I}_\beta(x_2)$.
 Using these expressions for the PDFs in the double-convolution of Eq.~(\ref{eq:pertexp}) one finds:
@@
 \end{equation}
-where the $C^{(s)}_{ij}$ matrix encodes the combination of PDFs that
+and the $C^{(s)}_{ij}$ matrix encodes the combination of PDFs that
 contribute to the $s$-th subprocess and $\hat{\sigma}^{(p)(s)}$
-is the corresponding hard cross section. Finally, defining:
+is the corresponding subprocess hard cross section. Finally, defining:
 \begin{equation}
@@
 
-\section{{\tt APFELgrid} manual}
-
-\section{Benchmark}
+\section{Performance benchmarks}
 
 In this section we discuss...
+
+\section{Conclusion}
 
@@
   \centering
+  \hspace*{-2.0cm}                                                           
   \includegraphics[scale=0.5]{plots/t0}
+\caption{\small Performance comparisons between {\tt APFELgrid} with
+  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
+  convolution time per point and process.}
+\label{fig:benchmark}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[ph]
+  \centering
+  \includegraphics[scale=0.5]{plots/t0a}
+\caption{\small Performance comparisons between {\tt APFELgrid} with
+  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
+  convolution time per point and process.}
+\label{fig:benchmark}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[ph]
+  \centering                                      
+  \includegraphics[scale=0.5]{plots/t0b}
 \caption{\small Performance comparisons between {\tt APFELgrid} with

### 2016-04-29 paper.tex
more edits

@@
 
-\section{The FastKernel method}\label{sec:FastKernel}
+\section{Interpolation tools for collider observables}\label{sec:FastKernel}
 
@@
 convolution of parton densities with a hard scattering cross-section. Consider for example the calculation of a general cross-section $pp\to X$ with a set of PDFs $\{f\}$:
-\begin{equation}\label{eq:pertexp}
+\begin{equation}\label{eq:hadconv}
   \sigma_{pp\to X} =
-  \sum_{i,j}\sum_{p=0} \int dx_1\,dx_2\,
-  \hat{\sigma}^{(p)}_{ij\to X}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)\,.
+  \sum_{s}\sum_{p} \int dx_1\,dx_2\,
+  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,.
 \end{equation}
-where $Q^2$ is the typical hard scale of the process, the indices
-$i,j$ sum over the active partonic species in the calculation, $p_{\rm LO}$ is the leading-order power of $\alpha_s$ for the process
-and $\hat{\sigma}^{(p)}_{ij\to X}$ the N$^p$LO contribution.
-
+where $Q^2$ is the typical hard scale of the process, the index
+$s$ sums over the active partonic subprocesses in the calculation, $p_{\rm LO}$ is the leading-order power of $\alpha_s$ for the process
+and $\hat{\sigma}^{(p)(s)}$ is the N$^p$LO cross-section for the partonic subprocess scattering $s\to X$. $F^{(s)}$ represents the 
+subprocess parton density:
+\begin{equation}\label{eq:APPLsubproc}
+  F^{(s)}(x_1,x_2, Q^2) =\sum_{i,j} C^{(s)}_{ij} 
+  f_i(x_{1},Q^2)f_j(x_{2},Q^2)\,,
+\end{equation}
+where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs that
+contribute to the $s$-th subprocess.
 The central observation of tools such as {\tt APPLgrid} and {\tt
@@
   the convolution via expansion over a set of interpolating functions, spanning $Q^2$ and the
-  two values of parton-$x$. For example one may represent the PDFs and $\alpha_S$ as
+  two values of parton-$x$.
+For example one may represent the subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and $\mathcal{I}_\beta(x_2)$ as:
 \begin{equation}\label{eq:interpolation}
 \begin{array}{c}
-\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,f_i(x_1,Q^2) f_j(x_2,Q^2)
+\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)
   =\\
@@
 \displaystyle \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
-  \,f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)
+  \,F^{(s)}_{\alpha\beta , \tau}
   \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
@@
 \end{equation}
-in terms of the Lagrange basis polynomials $\mathcal{I}_\tau(Q^2)$,
-$\mathcal{I}_\alpha(x_1)$, $\mathcal{I}_\beta(x_2)$.
-Using these expressions for the PDFs in the double-convolution of Eq.~(\ref{eq:pertexp}) one finds:
-\begin{equation}\label{eq:XsecPDFs}
-\begin{array}{rcl}
-  \displaystyle \sigma_{pp\to X} &=& \displaystyle 
-                                     \sum_{i,j}\sum_{p=0}^{N_\alpha} \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
-                                     \,f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)\\
-  \\
-                                 &\times& \displaystyle \mathcal{I}_\tau^{(n')}(Q^2)\int dx_1\,dx_2\,
-                                          \hat{\sigma}^{(p)}_{ij\to X}\,\mathcal{I}_\alpha^{(n_1)}(x_1)
-                                          \,\mathcal{I}_\beta^{(n_2)}(x_2)\,.
-\end{array}
-\end{equation}
+where we use the shorthand $F^{(s)}_{\alpha\beta ,\tau} = F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.
+Using these expressions in the double-convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an expression for the desired
+cross-section which depends upon the subprocess PDFs only though a simple product:
 
-Additionally one may take advantages of symmetries present in the matrix element to reduce the sum over active flavours $(i,j)$ in
-Eq.~(\ref{eq:hadconv}) to a reduced number of partonic subprocesses with a replacement such as
-\begin{equation}
-\sum_{i,j} \hat{\sigma}^{(p)}_{ij\to X}
-f_i(x_\alpha,Q_\tau^2)\,f_j(x_\beta,Q_\tau^2)\rightarrow
-\sum_{s}^{N_{\rm sub}} \hat{\sigma}^{(p)(s)} F^{(s)}_{\alpha\beta,\tau}
-\end{equation}
-where the $s$ index runs over the $N_{\rm sub}$ (process-specific)
-parton subprocess:
-\begin{equation}\label{eq:APPLsubproc}
-  F^{(s)}_{\alpha\beta,\tau} =\sum_{i,j} C^{(s)}_{ij} 
-  f_i(x_{\alpha},Q^2_\tau)f_j(x_{\beta},Q^2_\tau)\,,
-\end{equation}
-and the $C^{(s)}_{ij}$ matrix encodes the combination of PDFs that
-contribute to the $s$-th subprocess and $\hat{\sigma}^{(p)(s)}$
-is the corresponding subprocess hard cross section. Finally, defining:
-\begin{equation}
-  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau^{(n')}(Q^2)\int dx_1\,dx_2\,
-  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha^{(n_1)}(x_1)
-  \,\mathcal{I}_\beta^{(n_2)}(x_2)\,,
-\end{equation}
-one can write Eq.~(\ref{eq:XsecPDFs}) as follows:
 \begin{equation} \label{eq:applconv}
@@
 \end{equation}
+where
+\begin{equation}
+  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
+  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
+  \,\mathcal{I}_\beta(x_2)\,,
+\end{equation}
+consists of the convolution of the hard cross-section with the interpolating polynomials. This information may be stored in a precomputed look-up table.
+The final expression for the cross-section in Eqn.~\ref{eq:applconv} therefore being considerably more computationally efficient inside a fit than the direct
+performance of the double-convolution.
 
-Additionally, a number of tools, $e.g.$ {\tt
+\subsection{The FastKernel method}
+A number of tools, $e.g.$ {\tt
   APFEL}~\cite{Bertone:2013vaa} or {\tt HOPPET}~\cite{Salam:2008qg},

### 2016-04-29 paper.tex
more paper work

@@
 \\
-&=& \displaystyle \sum_{k,l}^{n_f}\sum_{\delta,\gamma}^{N_x}
+&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
 \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
@@
 \begin{equation}
-\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p^{N_\alpha}
-\sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta}
+\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
+\sum_{s} \sum_{\alpha,\beta}
 \sum_{\tau} 
@@
 
+\subsection{Properties of {\tt FK} tables}
+Some discussion here of the expected performance differences.
+Including precomputed sums over scale, perturbative order, no need to perform PDF evolution on the fly, smaller initial scale PDF basis, simple convolution structure that can be accellerated with SIMD/OpenMP.
+Explain reduced flexibility w.r.t $\alpha_S$ variations and scales, discuss increased storage space requirement but also reduced memory footprint.
+Might want to mention possibility for generating 'fast' grids as the interpolation accuracy is decoupled.
+
 \section{Performance benchmarks}
 
-In this section we discuss...
+The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eqn.~\ref{eq:FKTable} from weight grids supplied by {\tt APPLgrid} files.
+In this section we shall examine the performance differences between the two expressions for cross-section prediction Eqn.~\ref{eq:applconv} ({\tt APPLgrid}) and Eqn.~\ref{eq:FK} ({\tt FK}).
+For a comprehensive benchmark, we shall consider a wide range of processes including LHC and Tevatron electroweak vector boson production measurements, inclusive jet data and $t\bar{t}$ total cross-sections.
+Predictions are provided for a wide range of kinematics, for a total of 52 source {\tt APPLgrid} files roughly corresponding to the full collider dataset of the forthcoming NNPDF3.1 PDF determination.
 

### 2016-04-29 paper.tex
updating layout

@@
 \documentclass[preprint,12pt]{elsarticle}
+
+\textwidth=16.0cm \textheight=21.0cm 
+\topmargin 0cm \oddsidemargin 0cm 
+\setlength{\unitlength}{1mm}
 
@@
 \usepackage{amssymb}
+\usepackage{url,hyperref}
+
 %% The amsthm package provides extended theorem environments
@@
   fits of parton densities}
+\tnotetext[mytitlenote]{Preprints: OUTP-16-09P, CERN-TH-2016-103}
 
@@
 \address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
-\address[b]{TH Unit, CERN, CH-1211 Geneva 23, Switzerland}
+\address[b]{Theory Department, CERN, CH-1211 Geneva 23, Switzerland}
 
@@
 QCD; parton distribution functions; fast predictions.
-
 \end{keyword}
@@
 \end{frontmatter}
+
+ \newpage
+
+\noindent {\Large \textbf{Program Summary}}\\
 
@@
 
-%%\tableofcontents
+\tableofcontents
 
@@
 consists of the convolution of the hard cross-section with the interpolating polynomials. This information may be stored in a precomputed look-up table.
-The final expression for the cross-section in Eqn.~\ref{eq:applconv} therefore being considerably more computationally efficient inside a fit than the direct
+The final expression for the cross-section in Eq.~(\ref{eq:applconv}) therefore being considerably more computationally efficient inside a fit than the direct
 performance of the double-convolution.
@@
 
-The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eqn.~\ref{eq:FKTable} from weight grids supplied by {\tt APPLgrid} files.
-In this section we shall examine the performance differences between the two expressions for cross-section prediction Eqn.~\ref{eq:applconv} ({\tt APPLgrid}) and Eqn.~\ref{eq:FK} ({\tt FK}).
+The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt APPLgrid} files.
+In this section we shall examine the performance differences between the two expressions for cross-section prediction Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}).
 For a comprehensive benchmark, we shall consider a wide range of processes including LHC and Tevatron electroweak vector boson production measurements, inclusive jet data and $t\bar{t}$ total cross-sections.
@@
 
-\section{Conclusion}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ph]
-  \centering
-  \hspace*{-2.0cm}                                                           
-  \includegraphics[scale=0.5]{plots/t0}
-\caption{\small Performance comparisons between {\tt APFELgrid} with
-  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
-  convolution time per point and process.}
-\label{fig:benchmark}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ph]
+\begin{figure}[p]
   \centering
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ph]
+\begin{figure}[p]
   \centering                                      
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusion}
+
+\section*{Acknowledgements}
+
+We thanks the NNPDF Collaboration members for the support and
+motivation they gave us to write this work. V.~B. and N.~H. are
+supported by an European Research Council Starting Grant ``PDF4BSM''.
+S.~C. is supported by the HICCUP ERC Consolidator grant (614577).
 

### 2016-04-30 paper.tex
updating figures
updating section 3

@@
 % \biboptions{}
+\biboptions{numbers,sort&compress}
 
@@
   FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the hard
-cross sections, these packages rely on external codes to which they
+cross-sections, these packages rely on external codes to which they
 are interfaced by means of a suite of
@@
 the NNPDF global analyses~\cite{Ball:2010de}, was developed to address
-this problem. This method differs from the standard procedure \`{a} la {\tt APPLgrid} or {\tt fastNLO} in that it
+this problem. This method differs from the standard procedure \`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it
 maximises the amount of information that is precomputed prior to fitting such as to minimise the amount
 of operations to be performed during the fit. Specifically, the {\tt FastKernel} method relies on the precomputation
-of the partonic hard cross sections and DGLAP evolution kernels
+of the partonic hard cross-sections and DGLAP evolution kernels
 and their combination into a single look-up table, here called a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction for a given hadronic
@@
 
-In this paper we present the {\tt APFELgrid} package, a public implementation of the {\tt
-  FastKernel} method in which the hard partonic cross sections provided in an {\tt APPLgrid} look-up table are combined with the DGLAP
-evolution kernels provided by the {\tt APFEL}
-package~\cite{Bertone:2013vaa}.
+In this paper we present the {\tt APFELgrid} package, a public
+implementation of the {\tt FastKernel} method in which the hard
+partonic cross sections provided in an {\tt APPLgrid} look-up table
+are combined with the DGLAP evolution kernels provided by the {\tt
+  APFEL}
+package~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}.
 
 The outline of this paper is the following. In
-Sect.~\ref{sec:FastKernel} we present the tecnical details of the
-implemantation of the {\tt FastKernel} method.
+Sect.~\ref{sec:FastKernel} we present the technical details of the
+implementation of the {\tt FastKernel} method. Then, in
+Sect.~\ref{sec:benchmark} we illustrate timings and performance
+results of the {\tt APFELgrid} library and grids. Finally, in
+Sect.~\ref{sec:conclusion} we conclude and discuss possible future
+developments.
 
@@
 A number of tools, $e.g.$ {\tt
-  APFEL}~\cite{Bertone:2013vaa} or {\tt HOPPET}~\cite{Salam:2008qg},
-are available which perform PDF evolution via an analogous
-procedure. {\tt APFEL} in particular, is able to provide the so-called
-\textit{evolution operator}. The evolution operator, is computed
-assuming that the DGLAP evolution can be encoded in a single operator
-whose action on the intial scale PDFs proceduces the final scale
-PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
-on the $x$-space grid one can write:
+  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}
+or {\tt HOPPET}~\cite{Salam:2008qg}, are available which perform PDF
+evolution via an analogous procedure. {\tt APFEL} in particular, is
+able to provide the so-called \textit{evolution operator}. The
+evolution operator, is computed assuming that the DGLAP evolution can
+be encoded in a single operator whose action on the intial scale PDFs
+proceduces the final scale PDFs. Taking $Q_0$ as an initial scale and
+$Q_\tau$ as a final scale, on the $x$-space grid one can write:
 \begin{equation}\label{eq:fastPDFfinal_recalled}
@@
 \section{Performance benchmarks}
+\label{sec:benchmark}
 
-The {\tt APFELgrid} plugin generates {\tt FK} tables of the form expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt APPLgrid} files.
-In this section we shall examine the performance differences between the two expressions for cross-section prediction Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}).
-For a comprehensive benchmark, we shall consider a wide range of processes including LHC and Tevatron electroweak vector boson production measurements, inclusive jet data and $t\bar{t}$ total cross-sections.
-Predictions are provided for a wide range of kinematics, for a total of 52 source {\tt APPLgrid} files roughly corresponding to the full collider dataset of the forthcoming NNPDF3.1 PDF determination.
+The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
+expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
+  APPLgrid} files. In this section we shall examine the performance
+differences between the two expressions for cross-section prediction
+Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
+  FK}).
+%
+For a comprehensive benchmark, we consider here a wide range of
+processes including LHC and Tevatron electroweak vector boson
+production measurements, $t\bar{t}$ total cross-sections,
+double-differential Drell-Yan cross-sections and inclusive jet
+data. Predictions are provided for a wide range of kinematics, for a
+total of 52 source {\tt APPLgrid} files roughly corresponding to the
+full collider dataset of the forthcoming NNPDF3.1 PDF determination.
+
+In Fig.~\ref{fig:timings} we compare the convolution time per point
+between {\tt APFELgrid} and {\tt APPLgrid} for all the 52
+processes. We show timings for the {\tt APFELgrid} library in four
+different configurations: AVX-OpenMP 2x (2 CPUs), AVX SSE3, and
+standard double precision dot product convolution. When comparing the
+results we observe a great performance improvement for convolution
+time per point when using {\tt APFELgrid}. In particular, when
+comparing {\tt APFELgrid} AVX to {\tt APPLgrid} timings we obtain a
+minimum of x10 improvement for electroweak vector boson production and
+a maximum of x2000 for inclusive jet data. These results confirm the
+great usefulness of {\tt APFELgrid} in terms of performance for fits
+of PDFs.
+
+In Fig.~\ref{fig:performance} we provide information about the {\tt
+  APFELgrid} table generation, grid size and memory usage. In terms of
+generation time per point, we observe a timings from few milliseconds
+to 3.5 minutes per point. Such difference is almost due to the density
+of weights contained in the input {\tt APFELgrid} requiring a larger
+number of combination operations. When considering the grid size, we
+confirm, as expected, that in several processes the {\tt APPLgrid}
+files are smaller than {\tt APFELgrid} ones. This feature is justified
+by the lack of compression of {\tt APFELgrid} files. Finally, when
+measuring the resident set size memory used by the convolution
+programs, the amount of memory required by {\tt APFELgrid} is smaller
+than {\tt APPLgrid} for all processes considered here.
+
+In conclusion, the great performance benefits obtained with {\tt
+  APFELgrid} endorse the importance of this tool not only in PDF fits
+but also in any application requiring a large number of convolutions.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[p]
+\begin{figure}
   \centering
-  \includegraphics[scale=0.5]{plots/t0a}
+  \includegraphics[scale=0.6]{plots/t0a}
 \caption{\small Performance comparisons between {\tt APFELgrid} with
-  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
-  convolution time per point and process.}
-\label{fig:benchmark}
+  AVX-OpenMP 2x, AVX, SSE3, double precision convolution and {\tt
+    APPLgrid} convolution time per point and process.}
+\label{fig:timings}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[p]
+\begin{figure}
   \centering                                      
-  \includegraphics[scale=0.5]{plots/t0b}
-\caption{\small Performance comparisons between {\tt APFELgrid} with
-  AVX, SSE3 and double precision convolution and {\tt APPLgrid}
-  convolution time per point and process.}
-\label{fig:benchmark}
+  \includegraphics[scale=0.6]{plots/t0b}
+\caption{\small {\tt APFELgrid} generation time per point and
+  comparison to {\tt APPLgrid} for grid size and resident set size
+  (RSS).}
+\label{fig:performance}
 \end{figure}
@@
 \section{Conclusion}
+\label{sec:conclusion}
+
+In this work we presented {\tt APFELgrid}, an efficient convolution
+interface to hadronic observables for fits of parton densities.
 
@@
 %% References with bibTeX database:
-
+\section*{References}
 \bibliographystyle{elsarticle-num}
-\bibliography{paper.bib}
+\bibliography{paper}
 

### 2016-05-01 paper.tex
updating draft

@@
 \address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
-\address[b]{Theory Department, CERN, CH-1211 Geneva 23, Switzerland}
+\address[b]{Theoretical Physics Department, CERN, Geneva, Switzerland}
 
@@
 
-The {\tt FastKernel} method, developed in the framework of
-the NNPDF global analyses~\cite{Ball:2010de}, was developed to address
-this problem. This method differs from the standard procedure \`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it
-maximises the amount of information that is precomputed prior to fitting such as to minimise the amount
-of operations to be performed during the fit. Specifically, the {\tt FastKernel} method relies on the precomputation
-of the partonic hard cross-sections and DGLAP evolution kernels
-and their combination into a single look-up table, here called a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction for a given hadronic
-observable can be obtained by a simple matrix product between
-the respective {\tt FK} table and the initial scale PDFs.
+The {\tt FastKernel} method, developed in the framework of the NNPDF
+global analyses~\cite{Ball:2010de,Ball:2014uwa}, was developed to
+address this problem. This method differs from the standard procedure
+\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
+amount of information that is precomputed prior to fitting such as to
+minimise the amount of operations to be performed during the
+fit. Specifically, the {\tt FastKernel} method relies on the
+precomputation of the partonic hard cross-sections and DGLAP evolution
+kernels and their combination into a single look-up table, here called
+a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction
+for a given hadronic observable can be obtained by a simple matrix
+product between the respective {\tt FK} table and the initial scale
+PDFs.
 
@@
 processes including LHC and Tevatron electroweak vector boson
-production measurements, $t\bar{t}$ total cross-sections,
-double-differential Drell-Yan cross-sections and inclusive jet
-data. Predictions are provided for a wide range of kinematics, for a
-total of 52 source {\tt APPLgrid} files roughly corresponding to the
-full collider dataset of the forthcoming NNPDF3.1 PDF determination.
+production
+measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
+$t\bar{t}$ total
+cross-sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
+double-differential Drell-Yan
+cross-sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
+data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
+are provided for a wide range of kinematics, for a total of 52 source
+{\tt APPLgrid} files roughly corresponding to the full collider
+dataset of the forthcoming NNPDF3.1 PDF determination.
 
@@
 comparing {\tt APFELgrid} AVX to {\tt APPLgrid} timings we obtain a
-minimum of x10 improvement for electroweak vector boson production and
-a maximum of x2000 for inclusive jet data. These results confirm the
+minimum of 10x improvement for electroweak vector boson production and
+a maximum of 2000x for inclusive jet data. These results confirm the
 great usefulness of {\tt APFELgrid} in terms of performance for fits
@@
   APFELgrid} table generation, grid size and memory usage. In terms of
-generation time per point, we observe a timings from few milliseconds
-to 3.5 minutes per point. Such difference is almost due to the density
-of weights contained in the input {\tt APFELgrid} requiring a larger
-number of combination operations. When considering the grid size, we
-confirm, as expected, that in several processes the {\tt APPLgrid}
-files are smaller than {\tt APFELgrid} ones. This feature is justified
-by the lack of compression of {\tt APFELgrid} files. Finally, when
+generation time per point, we observe timings from few milliseconds to
+3.5 minutes per point. Such difference is almost due to the density of
+weights contained in the input {\tt APPLgrid} which requires a larger
+number of combination operations. When considering the grid size we
+confirm that in several processes the {\tt APPLgrid} files are smaller
+than {\tt APFELgrid} ones as expected. This feature is justified by
+the lack of compression of {\tt APFELgrid} files. Finally, when
 measuring the resident set size memory used by the convolution
 programs, the amount of memory required by {\tt APFELgrid} is smaller
-than {\tt APPLgrid} for all processes considered here.
+than {\tt APPLgrid} for all processes considered here by at least 75\%.
 

### 2016-05-02 paper.tex
first version of the conclusions

@@
 \begin{abstract}
-We present a new software package designed to aid the inclusion of hadronic
-observables into Parton Distribution Function (PDF) fits. The {\tt APFELgrid} package
-converts interpolated weight tables provided by {\tt APPLgrid} files into a more efficient
-format for PDF fitting by the pre-computation and combination with PDF and $\alpha_s$ evolution factors provided
-by {\tt APFEL}. This combination significantly reduces the number of operations required to perform the calculation
-of hadronic observables in PDF fits and simplifies the structure of the calculation into
-a readily optimised scalar product. We demonstrate that our technique
-can lead to a substantial speed improvement as compared to the
-existing interfaces without any reduction in numerical accuracy.
-
-%We present a new method conceived to ease the inclusion of hadronic
-%observables into parton distribution function (PDF) fits. Our
-%technique relies on the the same principles adopted by the existing
-%fast interfaces but implements a set of improvements aimed at
-%optimizing the performance in the context of PDF fits where many
-%iterations and many predictions for each iteration are usually
-%required. The main novelties are the precomputation of the PDF and
-%$\alpha_s$ evolution and the optimisation of the numerical convolution
-%between hard cross sections and PDFs. We demonstrate that our technique
-%can lead to a substantial speed improvement as compared to the
-%existing interfaces.
+  We present a new software package designed to aid the inclusion of
+  hadronic observables into Parton Distribution Function (PDF)
+  fits. The {\tt APFELgrid} package converts interpolated weight
+  tables provided by {\tt APPLgrid} files into a more efficient format
+  for PDF fitting by the pre-computation and combination with PDF and
+  $\alpha_s$ evolution factors provided by {\tt APFEL}. This
+  combination significantly reduces the number of operations required
+  to perform the calculation of hadronic observables in PDF fits and
+  simplifies the structure of the calculation into a readily optimised
+  scalar product. We demonstrate that our technique can lead to a
+  substantial speed improvement as compared to the existing interfaces
+  without any reduction in numerical accuracy.
+  % We present a new method conceived to ease the inclusion of
+  % hadronic observables into parton distribution function (PDF)
+  % fits. Our technique relies on the the same principles adopted by
+  % the existing fast interfaces but implements a set of improvements
+  % aimed at optimizing the performance in the context of PDF fits
+  % where many iterations and many predictions for each iteration are
+  % usually required. The main novelties are the precomputation of the
+  % PDF and $\alpha_s$ evolution and the optimisation of the numerical
+  % convolution between hard cross sections and PDFs. We demonstrate
+  % that our technique can lead to a substantial speed improvement as
+  % compared to the existing interfaces.
 \end{abstract}
@@
 
-The ability to perform fast and accurate cross-section predictions for measurements at
-the Large Hadron Collider (LHC) is a vital requirement for the most precise PDF determinations. LHC measurements have a unique
-capacity to shed light upon the dynamics of the proton and provide stringent constraints upon 
-the proton PDF. In order to make full use of current and future LHC data, efficient strategies
-for the computation of predictions for these observables must be deployed. Software packages performing predictions 
-for non-inclusive hadron-collider observables cannot be easily deployed in a PDF fit requiring rapid iteration due the time required to obtain accurate results (usually of the order of a few hours or more
-per data point). In order to overcome this limitation, the typical strategy adopted for fast cross-section prediction relies on the precomputation
-of the partonic hard cross sections in such a way that the numerical convolution with any set of PDFs can be quickly performed by means of interpolation techniques.
+The ability to perform fast and accurate cross-section predictions for
+measurements at the Large Hadron Collider (LHC) is a vital requirement
+for the most precise PDF determinations. LHC measurements have a
+unique capacity to shed light upon the dynamics of the proton and
+provide stringent constraints upon the proton PDF. In order to make
+full use of current and future LHC data, efficient strategies for the
+computation of predictions for these observables must be
+deployed. Software packages performing predictions for non-inclusive
+hadron-collider observables cannot be easily deployed in a PDF fit
+requiring rapid iteration due the time required to obtain accurate
+results (usually of the order of a few hours or more per data
+point). In order to overcome this limitation, the typical strategy
+adopted for fast cross-section prediction relies on the precomputation
+of the partonic hard cross sections in such a way that the numerical
+convolution with any set of PDFs can be quickly performed by means of
+interpolation techniques.
 
@@
   APPLgrid}~\cite{Carli:2010rw} and {\tt
-  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the hard
-cross-sections, these packages rely on external codes to which they
-are interfaced by means of a suite of
-functions allowing for the filling of PDF- and
-$\alpha_s$-independent look-up tables of cross-section weights. Codes
-such as {\tt MCFM}~\cite{Campbell:2010ff} and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to 
+  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the
+hard cross-sections, these packages rely on external codes to which
+they are interfaced by means of a suite of functions allowing for the
+filling of PDF- and $\alpha_s$-independent look-up tables of
+cross-section weights. Codes such as {\tt MCFM}~\cite{Campbell:2010ff}
+and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
 {\tt APPLgrid}/{\tt FastNLO} and more recently dedicated interfaces to
 automated general-purpose Monte Carlo (MC) event generators have been
-developed. The {\tt aMCfast} and {\tt MCgrid} codes
-have been specifically designed to extract the relevant information
-from the {\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
+developed. The {\tt aMCfast}~\cite{Bertone:2014zva} and {\tt
+  MCgrid}~\cite{DelDebbio:2013kxa} codes have been specifically
+designed to extract the relevant information from the {\tt
+  MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
   SHERPA}~\cite{Gleisberg:2008ta} event generators respectively.
 
-While these tools have proven to be extremely useful in the
-extraction of parton densities, the volume of experimental data
-made available by LHC collaborations for use in PDF fits is already
-stretching the capabilities of typical fitting technology. Presently a typical global PDF fit includes
-thousands of hadronic data points for which predictions have to be
-computed thousands of times during the minimisation process. As a
-consequence the ``standard'' tools, $i.e.$ {\tt APPLgrid} and {\tt
-  FastNLO}, might not be fast enough to meet the needs of modern
-PDF fitters. For this reason a high-performance tool tailored specifically 
-for PDF analyses is becoming increasingly important.
+While these tools have proven to be extremely useful in the extraction
+of parton densities, the volume of experimental data made available by
+LHC collaborations for use in PDF fits is already stretching the
+capabilities of typical fitting technology. Presently a typical global
+PDF fit includes thousands of hadronic data points for which
+predictions have to be computed thousands of times during the
+minimisation process. As a consequence the ``standard'' tools, $i.e.$
+{\tt APPLgrid} and {\tt FastNLO}, might not be fast enough to meet the
+needs of modern PDF fitters. For this reason a high-performance tool
+tailored specifically for PDF analyses is becoming increasingly
+important.
 
@@
 Hadron collider observables are typically computed in QCD by a double
-convolution of parton densities with a hard scattering cross-section. Consider for example the calculation of a general cross-section $pp\to X$ with a set of PDFs $\{f\}$:
+convolution of parton densities with a hard scattering
+cross-section. Consider for example the calculation of a general
+cross-section $pp\to X$ with a set of PDFs $\{f\}$:
 \begin{equation}\label{eq:hadconv}
@@
 \end{equation}
-where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs that
-contribute to the $s$-th subprocess.
-The central observation of tools such as {\tt APPLgrid} and {\tt
-  FastNLO} is that the PDF and $\alpha_S$ dependence may be factorised out of
-  the convolution via expansion over a set of interpolating functions, spanning $Q^2$ and the
-  two values of parton-$x$.
-For example one may represent the subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and $\mathcal{I}_\beta(x_2)$ as:
+where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs
+that contribute to the $s$-th subprocess.  The central observation of
+tools such as {\tt APPLgrid} and {\tt FastNLO} is that the PDF and
+$\alpha_S$ dependence may be factorised out of the convolution via
+expansion over a set of interpolating functions, spanning $Q^2$ and
+the two values of parton-$x$. For example one may represent the
+subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis
+polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and
+$\mathcal{I}_\beta(x_2)$ as:
 \begin{equation}\label{eq:interpolation}
@@
 \end{equation}
-where we use the shorthand $F^{(s)}_{\alpha\beta ,\tau} = F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.
-Using these expressions in the double-convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an expression for the desired
-cross-section which depends upon the subprocess PDFs only though a simple product:
+where we use the shorthand
+$F^{(s)}_{\alpha\beta ,\tau} = F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.
+Using these expressions in the double-convolution of
+Eq.~(\ref{eq:hadconv}) one can finally obtain an expression for the
+desired cross-section which depends upon the subprocess PDFs only
+though a simple product:
 
@@
 \end{equation}
-consists of the convolution of the hard cross-section with the interpolating polynomials. This information may be stored in a precomputed look-up table.
-The final expression for the cross-section in Eq.~(\ref{eq:applconv}) therefore being considerably more computationally efficient inside a fit than the direct
-performance of the double-convolution.
+consists of the convolution of the hard cross-section with the
+interpolating polynomials. This information may be stored in a
+precomputed look-up table.  The final expression for the cross-section
+in Eq.~(\ref{eq:applconv}) therefore being considerably more
+computationally efficient inside a fit than the direct performance of
+the double-convolution.
 
 \subsection{The FastKernel method}
+
 A number of tools, $e.g.$ {\tt
@@
 \end{equation}
-where the operator $A$ is the solution of the DGLAP evolution
-equation with baundary condition:
+where the operator $A$ is the solution of the DGLAP evolution equation
+with baundary condition:
 \begin{equation}
@@
 \subsection{Properties of {\tt FK} tables}
-Some discussion here of the expected performance differences.
-Including precomputed sums over scale, perturbative order, no need to perform PDF evolution on the fly, smaller initial scale PDF basis, simple convolution structure that can be accellerated with SIMD/OpenMP.
-Explain reduced flexibility w.r.t $\alpha_S$ variations and scales, discuss increased storage space requirement but also reduced memory footprint.
-Might want to mention possibility for generating 'fast' grids as the interpolation accuracy is decoupled.
+Some discussion here of the expected performance
+differences. Including precomputed sums over scale, perturbative
+order, no need to perform PDF evolution on the fly, smaller initial
+scale PDF basis, simple convolution structure that can be accellerated
+with SIMD/OpenMP.  Explain reduced flexibility w.r.t $\alpha_S$
+variations and scales, discuss increased storage space requirement but
+also reduced memory footprint.  Might want to mention possibility for
+generating 'fast' grids as the interpolation accuracy is decoupled.
 
@@
 programs, the amount of memory required by {\tt APFELgrid} is smaller
-than {\tt APPLgrid} for all processes considered here by at least 75\%.
+than {\tt APPLgrid} for all processes considered here by at least
+75\%.
 
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
-In this work we presented {\tt APFELgrid}, an efficient convolution
-interface to hadronic observables for fits of parton densities.
+In this work we have presented {\tt APFELgrid}, an efficient
+convolution interface to hadronic observables for fits of parton
+densities. We have demonstrated that employing the so-called {\tt
+  FastKernel} method, it is possible to convert an {\tt APPLgrid}
+weight table into a derived format, referred to as {\tt FKtable},
+which includes PDF and $\alpha_s$ evolution as provided by {\tt
+  APFEL}.  Such a procedure allows one to obtain a very
+computationally-efficient expression for the calculation of hadronic
+cross sections, in terms of only the initial-scale PDFs and summing
+only over the initial scale interpolating grid and the incoming parton
+flavors. In addition, the simple structure of the resulting
+convolution makes {\tt FKtables} particularly suitable to make
+efficient use of acceleration and parallelization techniques like {\tt
+  SIMD} and {\tt OpenMD}.
+
+We have shown that the numerical convolution of an {\tt APFELgrid}
+tables is typically much faster than the corresponding {\tt
+  APPLgrid}. In particular, the {\tt FastKernel} method provides a
+great perfomance improvement with respect to {\tt APPLgrid} which is
+further enhanced by the employment of any of the acceleration
+techniques available in {\tt APFELgrid}. Finally, while the size of an
+{\tt FKtable} is often larger than the corresponding {\tt APFELgrid}
+file, we have shown that the resident set size memory required by {\tt
+  APFELgrid} is typically smaller than {\tt APPLgrid} by at least
+75\%.
+
+The subtantial speed improvement with respect to {\tt APPLgrid} in
+association with the reduction of the resident set size memory, makes
+the {\tt APFELgrid} code a valuable tool for modern PDF fits which
+usually incluse thousands of hadronic data points and that thus
+heavily rely on fast interfaces. We stress that {\tt APFELgrid} is not
+a replacement for {\tt APPLgrid}, on which it depends, but it is
+rather a specialized tool conceived to ease the inclusion of hadronic
+data into PDF fits. In fact, the price of a better performance is the
+fact that, by themselves, {\tt FKtables} cannot perform any parameter
+variation other than that of PDFs. For example, variations of
+$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.
+
+The {\tt APFELgrid} code is publicly available from the webpage:
+\begin{center}
+\url{https://github.com/nhartland/APFELgrid}
+\end{center}
+where intructions to install and run codes are also provided.
 

### 2016-05-03 paper.tex
Sec 3 edits

@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 order, no need to perform PDF evolution on the fly, smaller initial
-scale PDF basis, simple convolution structure that can be accellerated
+scale PDF basis, simple convolution structure that can be accelerated
 with SIMD/OpenMP.  Explain reduced flexibility w.r.t $\alpha_S$
@@
 differences between the two expressions for cross-section prediction
-Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
-  FK}).
+Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}).
 %
-For a comprehensive benchmark, we consider here a wide range of
+To provide a comprehensive benchmark, we consider here a wide range of
 processes including LHC and Tevatron electroweak vector boson
@@
 data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
-are provided for a wide range of kinematics, for a total of 52 source
-{\tt APPLgrid} files roughly corresponding to the full collider
-dataset of the forthcoming NNPDF3.1 PDF determination.
+are performed over a wide range of kinematics, for a total of 52 source
+{\tt APPLgrid} files corresponding to the majority of available LHC and Tevatron datasets of applicability to PDF determination.
+While the source {\tt APPLgrid} files have varying grid densities in $x$ and $Q^2$, for the purposes of comparison the corresponding {\tt FK} tables are produced
+consistently with 30 points in $x$, and at an initial scale below the charm threshold, therefore with seven active partonic species.
 
-In Fig.~\ref{fig:timings} we compare the convolution time per point
-between {\tt APFELgrid} and {\tt APPLgrid} for all the 52
-processes. We show timings for the {\tt APFELgrid} library in four
-different configurations: AVX-OpenMP 2x (2 CPUs), AVX SSE3, and
-standard double precision dot product convolution. When comparing the
-results we observe a great performance improvement for convolution
-time per point when using {\tt APFELgrid}. In particular, when
-comparing {\tt APFELgrid} AVX to {\tt APPLgrid} timings we obtain a
-minimum of 10x improvement for electroweak vector boson production and
-a maximum of 2000x for inclusive jet data. These results confirm the
-great usefulness of {\tt APFELgrid} in terms of performance for fits
-of PDFs.
+In Fig.~\ref{fig:timings} we compare the average time taken per datapoint
+for the {\tt FK} and {\tt APPLgrid} calculations, for all of the 52
+tables. We show timings for the {\tt FK} calculation in four
+different configurations: AVX-OpenMP 2x (2 CPU cores), AVX, SSE3 and
+standard double precision dot product. Due to the inherent structural differences between the {\tt FK} and {\tt APPLgrid} procedures,
+results from the {\tt FK} calculation are systematically faster than those from {\tt APPLgrids}.
+In particular, when comparing {\tt FK} AVX to {\tt APPLgrid} timings we obtain minimally a factor
+of ten improvement in speed for electroweak vector boson production and
+a maximum factor of 2000 improvement in predictions for inclusive jet data. Across all processes
+and kinematic regions we observe significant performance improvements from using the {\tt FK} calculation
+even in the maximally compatible, non-SIMD case.
 
-In Fig.~\ref{fig:performance} we provide information about the {\tt
-  APFELgrid} table generation, grid size and memory usage. In terms of
-generation time per point, we observe timings from few milliseconds to
-3.5 minutes per point. Such difference is almost due to the density of
-weights contained in the input {\tt APPLgrid} which requires a larger
-number of combination operations. When considering the grid size we
-confirm that in several processes the {\tt APPLgrid} files are smaller
-than {\tt APFELgrid} ones as expected. This feature is justified by
-the lack of compression of {\tt APFELgrid} files. Finally, when
-measuring the resident set size memory used by the convolution
-programs, the amount of memory required by {\tt APFELgrid} is smaller
-than {\tt APPLgrid} for all processes considered here by at least
-75\%.
+While sheer computational speed is typically the primary consideration when computing observables
+in a PDF fit, other factors such as table size in the filesystem and memory, along with the computational cost of
+pre-computing {\tt FK} tables must be considered. Indeed, the computation of the {\tt FK} table in Eqn.~\ref{eq:FKTable}
+requires a great deal of operations which can be time consuming, particuarly in the case of source {\tt APPLgrids} with very high
+interpolation precision.
 
-In conclusion, the great performance benefits obtained with {\tt
-  APFELgrid} endorse the importance of this tool not only in PDF fits
-but also in any application requiring a large number of convolutions.
+In Fig.~\ref{fig:performance} we examine the {\tt FK} table generation time with {\tt APFELgrid}, {\tt FK} table file size and memory usage of the {\tt FK} tables arising from the same source {\tt APPLgrids} as discussed in Fig.~\ref{fig:timings}.
+When examining the table generation time per point, we observe timings from a few milliseconds to
+3.5 minutes per point, with differences arising from the varying grid densities used in the input {\tt APPLgrid} files. 
+In terms of the grid size on disc, {\tt FK} tables are typically larger than their corresponding {\tt APPLgrids}, primarily as the {\tt FK} file format is encoding in plain text for compatibility whereas {\tt APPLgrids} are expressed in binary as {\tt ROOT} files.
+However when
+measuring the in-memory resident set size used by the both convolution procedures, the amount of memory required by {\tt FK} tables is systematically less than {\tt APPLgrid} for all processes considered here by at least 75\%.
+
+%In conclusion, the great performance benefits obtained with {\tt
+%APFELgrid} endorse the importance of this tool not only in PDF fits
+%but also in any application requiring a large number of convolutions.
 

### 2016-05-03 paper.tex
features of fk tables

@@
 \end{equation}
-is referred to here as an {\tt FKtable}, and combines the information
+is referred to here as an {\tt FK} table, and combines the information
 stored in {\tt APPLgrid}-style interpolated weight grids with
-analogously interpolated DGLAP evolution operators. Such a combination
+analogously interpolated DGLAP evolution operators. This combination
 enables for a maximally computationally-efficient expression for the
-calculation of observables at hadron colliders.
+calculation of observables at hadron colliders under PDF variation.
 
-\subsection{Properties of {\tt FK} tables}
-Some discussion here of the expected performance
-differences. Including precomputed sums over scale, perturbative
-order, no need to perform PDF evolution on the fly, smaller initial
-scale PDF basis, simple convolution structure that can be accelerated
-with SIMD/OpenMP.  Explain reduced flexibility w.r.t $\alpha_S$
-variations and scales, discuss increased storage space requirement but
-also reduced memory footprint.  Might want to mention possibility for
-generating 'fast' grids as the interpolation accuracy is decoupled.
+\subsection{Features and limitations of {\tt FK} tables}
+The {\tt FK} product of Eqn.~\ref{eq:FK} differs with respect to the product in Eqn.~\ref{eq:applconv} in several notable ways. Firstly 
+the typical {\tt APPLgrid} or {\tt FastNLO} products require PDFs at a general scale $Q^2$, which implies performing PDF evolution e.g Eqn.~\ref{eq:fastPDFfinal_recalled} for every variation of the PDFs at the input scale during the fit.
+However in the {\tt FK} convolution this evolution
+is pre-cached at the stage of {\tt FK} table generation, requiring only initial-scale PDFs as an input at fitting time. This pre-caching of the evolution
+also removes the need for performing the sums over hard scale and perturbative order during the fit, further reducing the number of operations required.
+The {\tt FK} convolution also benefits from the typically reduced number of active partonic modes at the fitting scale, with the sum over flavours in Eqn.~\ref{eq:FK} being limited to those directly used in the fit.
+Having reduced the calculation to such a simple form, it is also possible to straightforwardly
+apply standard and widely-available computational tools such as multi-threading through e.g OpenMP or Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to further reduce computational cost.
 
+While these features provide significant performance enhancements, the {\tt FK } table format is not suitable as a complete replacement for tools
+such as {\tt APPLgrid}.
+The pre-computation of the PDF evolution necessarily means that all theory parameters such as perturbative order, strong coupling and factorization/renormalization scales are inextricably embedded in each {\tt FK} table.
+In order to perform PDF fits including variations of these parameters, multiple {\tt FK } tables must be computed, each with different theory settings.
+While performing such a re-calculation directly from Monte Carlo codes
+would be exceptionally time consuming, the data representation in {\tt APPLgrid} files allows for an efficient combination with varying theory parameters.
+
+%\section{The {\tt APFELgrid} package}
+% **** Not sure if this sort of section would be neccesary ***
+%The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
+%expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
+%  APPLgrid} files. It is supplied as a set of {\tt C++} routines designed to
+%  supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
+%  capabilities. The plugin has a very light interface for performing the {\tt FK} combination,
+%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
+%  including the structure of the initial scale $x$ grid.
+%*** Would now discuss e.g SIMD implementation etc in advance of looking at the results in next section ***
 \section{Performance benchmarks}
@@
 
-The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
-expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
-  APPLgrid} files. In this section we shall examine the performance
+In this section we shall examine the performance
 differences between the two expressions for cross-section prediction
@@
 convolution makes {\tt FKtables} particularly suitable to make
-efficient use of acceleration and parallelization techniques like {\tt
-  SIMD} and {\tt OpenMD}.
+efficient use of acceleration and parallelisation techniques like {\tt
+  SIMD} and {\tt OpenMP}.
 
@@
   APPLgrid}. In particular, the {\tt FastKernel} method provides a
-great perfomance improvement with respect to {\tt APPLgrid} which is
+great performance improvement with respect to {\tt APPLgrid} which is
 further enhanced by the employment of any of the acceleration
@@
 
-The subtantial speed improvement with respect to {\tt APPLgrid} in
+The substantial speed improvement with respect to {\tt APPLgrid} in
 association with the reduction of the resident set size memory, makes
@@
 a replacement for {\tt APPLgrid}, on which it depends, but it is
-rather a specialized tool conceived to ease the inclusion of hadronic
+rather a specialised tool conceived to ease the inclusion of hadronic
 data into PDF fits. In fact, the price of a better performance is the

### 2016-05-03 paper.tex
conclusions pass

@@
 
-%\section{The {\tt APFELgrid} package}
-% **** Not sure if this sort of section would be neccesary ***
-%The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
-%expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
-%  APPLgrid} files. It is supplied as a set of {\tt C++} routines designed to
-%  supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
-%  capabilities. The plugin has a very light interface for performing the {\tt FK} combination,
-%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
-%  including the structure of the initial scale $x$ grid.
-%*** Would now discuss e.g SIMD implementation etc in advance of looking at the results in next section ***
 \section{Performance benchmarks}
@@
 
-In this work we have presented {\tt APFELgrid}, an efficient
-convolution interface to hadronic observables for fits of parton
-densities. We have demonstrated that employing the so-called {\tt
+In this work we have demonstrated that by employing the so-called {\tt
   FastKernel} method, it is possible to convert an {\tt APPLgrid}
-weight table into a derived format, referred to as {\tt FKtable},
-which includes PDF and $\alpha_s$ evolution as provided by {\tt
-  APFEL}.  Such a procedure allows one to obtain a very
-computationally-efficient expression for the calculation of hadronic
+weight table into a derived format, referred to as {\tt FK} table,
+which includes PDF and $\alpha_s$ evolution. This procedure has been 
+implemented in the {\tt APFELgrid} package, supplied as a set of {\tt C++} routines designed to
+supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
+capabilities. 
+%The plugin has a very light interface for performing the {\tt FK} combination,
+%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
+%  including the structure of the initial scale $x$ grid.
+The {\tt APFELgrid} package allows one to obtain a 
+computationally efficient expression for the calculation of hadronic
 cross sections, in terms of only the initial-scale PDFs and summing
-only over the initial scale interpolating grid and the incoming parton
-flavors. In addition, the simple structure of the resulting
-convolution makes {\tt FKtables} particularly suitable to make
-efficient use of acceleration and parallelisation techniques like {\tt
+only over the initial scale interpolating $x$-grid and the incoming parton
+flavors. Additionally the simple structure of the resulting
+convolution makes {\tt FK} tables particularly suitable to make
+efficient use of computational tools such as {\tt
   SIMD} and {\tt OpenMP}.
 
-We have shown that the numerical convolution of an {\tt APFELgrid}
-tables is typically much faster than the corresponding {\tt
-  APPLgrid}. In particular, the {\tt FastKernel} method provides a
-great performance improvement with respect to {\tt APPLgrid} which is
-further enhanced by the employment of any of the acceleration
-techniques available in {\tt APFELgrid}. Finally, while the size of an
-{\tt FKtable} is often larger than the corresponding {\tt APFELgrid}
-file, we have shown that the resident set size memory required by {\tt
-  APFELgrid} is typically smaller than {\tt APPLgrid} by at least
-75\%.
+We have shown that in several practical examples the numerical evaluation of an {\tt FK}
+product is much faster than the corresponding {\tt APPLgrid} product, even in the case where
+neither SIMD or multi-threading are applied. {\tt FK} tables are supplied in a simple plain-text 
+format in order to simplify the construction of user interfaces, and therefore typically are often larger 
+than the corresponding {\tt APFELgrid} file. However we have shown that the in-memory resident set size occupied by {\tt
+  FK} tables is typically smaller than that requred by {\tt APPLgrid} by at least 75\%.
 
-The substantial speed improvement with respect to {\tt APPLgrid} in
-association with the reduction of the resident set size memory, makes
-the {\tt APFELgrid} code a valuable tool for modern PDF fits which
-usually incluse thousands of hadronic data points and that thus
-heavily rely on fast interfaces. We stress that {\tt APFELgrid} is not
-a replacement for {\tt APPLgrid}, on which it depends, but it is
-rather a specialised tool conceived to ease the inclusion of hadronic
-data into PDF fits. In fact, the price of a better performance is the
-fact that, by themselves, {\tt FKtables} cannot perform any parameter
-variation other than that of PDFs. For example, variations of
-$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.
+The substantial speed improvement of {\tt FK} tables with respect to {\tt APPLgrid} in
+association with the reduction in memory footprint, makes
+the {\tt APFELgrid} code a valuable tool for modern PDF fits including large collider datasets and thus
+rely heavily upon fast calculations.
 
-The {\tt APFELgrid} code is publicly available from the webpage:
+%We stress that {\tt APFELgrid} is not
+%a replacement for {\tt APPLgrid}, on which it depends, but it is
+%rather a specialised tool conceived to ease the inclusion of hadronic
+%data into PDF fits. In fact, the price of a better performance is the
+%fact that, by themselves, {\tt FKtables} cannot perform any parameter
+%variation other than that of PDFs. For example, variations of
+%$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.
+
+The {\tt APFELgrid} package and associated documentation are publicly available from the webpage:
 \begin{center}
@@
 \end{center}
-where intructions to install and run codes are also provided.
 

### 2016-05-03 paper.tex
sc pass

@@
 results of the {\tt APFELgrid} library and grids. Finally, in
-Sect.~\ref{sec:conclusion} we conclude and discuss possible future
-developments.
+Sect.~\ref{sec:conclusion} we summarise the results discussed in this
+work.
 
@@
 \subsection{Features and limitations of {\tt FK} tables}
-The {\tt FK} product of Eqn.~\ref{eq:FK} differs with respect to the product in Eqn.~\ref{eq:applconv} in several notable ways. Firstly 
-the typical {\tt APPLgrid} or {\tt FastNLO} products require PDFs at a general scale $Q^2$, which implies performing PDF evolution e.g Eqn.~\ref{eq:fastPDFfinal_recalled} for every variation of the PDFs at the input scale during the fit.
-However in the {\tt FK} convolution this evolution
-is pre-cached at the stage of {\tt FK} table generation, requiring only initial-scale PDFs as an input at fitting time. This pre-caching of the evolution
-also removes the need for performing the sums over hard scale and perturbative order during the fit, further reducing the number of operations required.
-The {\tt FK} convolution also benefits from the typically reduced number of active partonic modes at the fitting scale, with the sum over flavours in Eqn.~\ref{eq:FK} being limited to those directly used in the fit.
-Having reduced the calculation to such a simple form, it is also possible to straightforwardly
-apply standard and widely-available computational tools such as multi-threading through e.g OpenMP or Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to further reduce computational cost.
 
-While these features provide significant performance enhancements, the {\tt FK } table format is not suitable as a complete replacement for tools
-such as {\tt APPLgrid}.
-The pre-computation of the PDF evolution necessarily means that all theory parameters such as perturbative order, strong coupling and factorization/renormalization scales are inextricably embedded in each {\tt FK} table.
-In order to perform PDF fits including variations of these parameters, multiple {\tt FK } tables must be computed, each with different theory settings.
-While performing such a re-calculation directly from Monte Carlo codes
-would be exceptionally time consuming, the data representation in {\tt APPLgrid} files allows for an efficient combination with varying theory parameters.
+The {\tt FK} product of Eq.~(\ref{eq:FK}) differs with respect to the
+product in Eq.~(\ref{eq:applconv}) in several notable ways. Firstly
+the typical {\tt APPLgrid} or {\tt FastNLO} products require PDFs at a
+general scale $Q^2$, which implies performing PDF evolution {\it e.g.}
+Eq.~(\ref{eq:fastPDFfinal_recalled}) for every variation of the PDFs
+at the input scale during the fit. However in the {\tt FK} convolution
+this evolution is pre-cached at the stage of {\tt FK} table
+generation, requiring only initial-scale PDFs as an input at fitting
+time. This pre-caching of the evolution also removes the need for
+performing the sums over hard scale and perturbative order during the
+fit, further reducing the number of operations required.  The {\tt FK}
+convolution also benefits from the typically reduced number of active
+partonic modes at the fitting scale, with the sum over flavours in
+Eq.~(\ref{eq:FK}) being limited to those directly used in the
+fit. Having reduced the calculation to such a simple form, it is also
+possible to straightforwardly apply standard and widely-available
+computational tools such as multi-threading through e.g OpenMP or
+Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to
+further reduce computational cost.
+
+While these features provide significant performance enhancements, the
+{\tt FK} table format is not suitable as a complete replacement for
+tools such as {\tt APPLgrid}. The pre-computation of the PDF evolution
+necessarily means that all theory parameters such as perturbative
+order, strong coupling and factorization/renormalization scales are
+inextricably embedded in each {\tt FK} table. In order to perform PDF
+fits including variations of these parameters, multiple {\tt FK}
+tables must be computed, each with different theory settings. While
+performing such a re-calculation directly from Monte Carlo codes would
+be exceptionally time consuming, the data representation in {\tt
+  APPLgrid} files allows for an efficient combination with varying
+theory parameters.
 
@@
 
-In this section we shall examine the performance
-differences between the two expressions for cross-section prediction
-Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}).
+In this section we shall examine the performance differences between
+the two expressions for cross-section prediction
+Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
+  FK}).
 %
@@
 data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
-are performed over a wide range of kinematics, for a total of 52 source
-{\tt APPLgrid} files corresponding to the majority of available LHC and Tevatron datasets of applicability to PDF determination.
-While the source {\tt APPLgrid} files have varying grid densities in $x$ and $Q^2$, for the purposes of comparison the corresponding {\tt FK} tables are produced
-consistently with 30 points in $x$, and at an initial scale below the charm threshold, therefore with seven active partonic species.
+are performed over a wide range of kinematics, for a total of 52
+source {\tt APPLgrid} files corresponding to the majority of available
+LHC and Tevatron datasets of applicability to PDF determination.
+While the source {\tt APPLgrid} files have varying grid densities in
+$x$ and $Q^2$, for the purposes of comparison the corresponding {\tt
+  FK} tables are produced consistently with 30 points in $x$, and at
+an initial scale below the charm threshold, therefore with seven
+active partonic species.
 
-In Fig.~\ref{fig:timings} we compare the average time taken per datapoint
-for the {\tt FK} and {\tt APPLgrid} calculations, for all of the 52
-tables. We show timings for the {\tt FK} calculation in four
+In Fig.~\ref{fig:timings} we compare the average time taken per
+datapoint for the {\tt FK} and {\tt APPLgrid} calculations, for all of
+the 52 tables. We show timings for the {\tt FK} calculation in four
 different configurations: AVX-OpenMP 2x (2 CPU cores), AVX, SSE3 and
-standard double precision dot product. Due to the inherent structural differences between the {\tt FK} and {\tt APPLgrid} procedures,
-results from the {\tt FK} calculation are systematically faster than those from {\tt APPLgrids}.
-In particular, when comparing {\tt FK} AVX to {\tt APPLgrid} timings we obtain minimally a factor
-of ten improvement in speed for electroweak vector boson production and
-a maximum factor of 2000 improvement in predictions for inclusive jet data. Across all processes
-and kinematic regions we observe significant performance improvements from using the {\tt FK} calculation
-even in the maximally compatible, non-SIMD case.
+standard double precision dot product. Due to the inherent structural
+differences between the {\tt FK} and {\tt APPLgrid} procedures,
+results from the {\tt FK} calculation are systematically faster than
+those from {\tt APPLgrids}.  In particular, when comparing {\tt FK}
+AVX to {\tt APPLgrid} timings we obtain minimally a factor of ten
+improvement in speed for electroweak vector boson production and a
+maximum factor of 2000 improvement in predictions for inclusive jet
+data. Across all processes and kinematic regions we observe
+significant performance improvements from using the {\tt FK}
+calculation even in the maximally compatible, non-SIMD case.
 
-While sheer computational speed is typically the primary consideration when computing observables
-in a PDF fit, other factors such as table size in the filesystem and memory, along with the computational cost of
-pre-computing {\tt FK} tables must be considered. Indeed, the computation of the {\tt FK} table in Eqn.~\ref{eq:FKTable}
-requires a great deal of operations which can be time consuming, particuarly in the case of source {\tt APPLgrids} with very high
-interpolation precision.
+While sheer computational speed is typically the primary consideration
+when computing observables in a PDF fit, other factors such as table
+size in the filesystem and memory, along with the computational cost
+of pre-computing {\tt FK} tables must be considered. Indeed, the
+computation of the {\tt FK} table in Eq.~(\ref{eq:FKTable}) requires a
+great deal of operations which can be time consuming, particuarly in
+the case of source {\tt APPLgrids} with very high interpolation
+precision.
 
-In Fig.~\ref{fig:performance} we examine the {\tt FK} table generation time with {\tt APFELgrid}, {\tt FK} table file size and memory usage of the {\tt FK} tables arising from the same source {\tt APPLgrids} as discussed in Fig.~\ref{fig:timings}.
-When examining the table generation time per point, we observe timings from a few milliseconds to
-3.5 minutes per point, with differences arising from the varying grid densities used in the input {\tt APPLgrid} files. 
-In terms of the grid size on disc, {\tt FK} tables are typically larger than their corresponding {\tt APPLgrids}, primarily as the {\tt FK} file format is encoding in plain text for compatibility whereas {\tt APPLgrids} are expressed in binary as {\tt ROOT} files.
-However when
-measuring the in-memory resident set size used by the both convolution procedures, the amount of memory required by {\tt FK} tables is systematically less than {\tt APPLgrid} for all processes considered here by at least 75\%.
+In Fig.~\ref{fig:performance} we examine the {\tt FK} table generation
+time with {\tt APFELgrid}, {\tt FK} table file size and memory usage
+of the {\tt FK} tables arising from the same source {\tt APPLgrids} as
+discussed in Fig.~\ref{fig:timings}.  When examining the table
+generation time per point, we observe timings from a few milliseconds
+to 3.5 minutes per point, with differences arising from the varying
+grid densities used in the input {\tt APPLgrid} files.  In terms of
+the grid size on disk, {\tt FK} tables are typically larger than their
+corresponding {\tt APPLgrids}, primarily as the {\tt FK} file format
+is encoding in plain text for compatibility whereas {\tt APPLgrids}
+are expressed in binary as {\tt ROOT} files. However when measuring
+the in-memory resident set size used by the both convolution
+procedures, the amount of memory required by {\tt FK} tables is
+systematically less than {\tt APPLgrid} for all processes considered
+here by at least 75\%.
 
@@
 \caption{\small {\tt APFELgrid} generation time per point and
-  comparison to {\tt APPLgrid} for grid size and resident set size
+  comparison to {\tt APPLgrid} for grid size on disk and resident set size
   (RSS).}
@@
 which includes PDF and $\alpha_s$ evolution. This procedure has been
-implemented in the {\tt APFELgrid} package, supplied as a set of {\tt C++} routines designed to
-supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
-capabilities. 
+implemented in the {\tt APFELgrid} package, supplied as a set of {\tt
+  C++} routines designed to supplement the PDF evolution library {\tt
+  APFEL} with {\tt FK} table generation capabilities.
 %The plugin has a very light interface for performing the {\tt FK} combination,
@@
 %  including the structure of the initial scale $x$ grid.
-The {\tt APFELgrid} package allows one to obtain a 
-computationally efficient expression for the calculation of hadronic
-cross sections, in terms of only the initial-scale PDFs and summing
-only over the initial scale interpolating $x$-grid and the incoming parton
+The {\tt APFELgrid} package allows one to obtain a computationally
+efficient expression for the calculation of hadronic cross sections,
+in terms of only the initial-scale PDFs and summing only over the
+initial scale interpolating $x$-grid and the incoming parton
 flavors. Additionally the simple structure of the resulting
 convolution makes {\tt FK} tables particularly suitable to make
-efficient use of computational tools such as {\tt
-  SIMD} and {\tt OpenMP}.
+efficient use of computational tools such as {\tt SIMD} and {\tt
+  OpenMP}.
 
-We have shown that in several practical examples the numerical evaluation of an {\tt FK}
-product is much faster than the corresponding {\tt APPLgrid} product, even in the case where
-neither SIMD or multi-threading are applied. {\tt FK} tables are supplied in a simple plain-text 
-format in order to simplify the construction of user interfaces, and therefore typically are often larger 
-than the corresponding {\tt APFELgrid} file. However we have shown that the in-memory resident set size occupied by {\tt
-  FK} tables is typically smaller than that requred by {\tt APPLgrid} by at least 75\%.
+We have shown that in several practical examples the numerical
+evaluation of an {\tt FK} product is much faster than the
+corresponding {\tt APPLgrid} product, even in the case where neither
+SIMD or multi-threading are applied. {\tt FK} tables are supplied in a
+simple plain-text format in order to simplify the construction of user
+interfaces, and therefore typically are often larger than the
+corresponding {\tt APPLgrid} file. However we have shown that the
+in-memory resident set size occupied by {\tt FK} tables is typically
+smaller than that required by {\tt APPLgrid} by at least 75\%.
 
-The substantial speed improvement of {\tt FK} tables with respect to {\tt APPLgrid} in
-association with the reduction in memory footprint, makes
-the {\tt APFELgrid} code a valuable tool for modern PDF fits including large collider datasets and thus
-rely heavily upon fast calculations.
+The substantial speed improvement of {\tt FK} tables with respect to
+{\tt APPLgrid} in association with the reduction in memory footprint,
+makes the {\tt APFELgrid} code a valuable tool for modern PDF fits
+including large collider datasets and thus rely heavily upon fast
+calculations.
 

### 2016-05-04 paper.tex
edits

@@
 {\em Solution method:}\\
-  Combination of interpolated weight grids from {\tt APPLgrid} and evolution factors from {\tt APFEL} into efficient {\tt FastKernel} tables.\\
+  Combination of interpolated weight grids from {\tt APPLgrid} files and evolution factors from {\tt APFEL} into efficient {\tt FastKernel} tables.\\
 {\em Running time:} varying\\
@@
 and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
-{\tt APPLgrid}/{\tt FastNLO} and more recently dedicated interfaces to
-automated general-purpose Monte Carlo (MC) event generators have been
-developed. The {\tt aMCfast}~\cite{Bertone:2014zva} and {\tt
+{\tt APPLgrid}/{\tt FastNLO} and more recently de\-di\-ca\-ted
+interfaces to automated general-purpose Monte Carlo (MC) event
+generators have been developed. The {\tt
+  aMCfast}~\cite{Bertone:2014zva} and {\tt
   MCgrid}~\cite{DelDebbio:2013kxa} codes have been specifically
@@
 LHC collaborations for use in PDF fits is already stretching the
-capabilities of typical fitting technology. Presently a typical global
-PDF fit includes thousands of hadronic data points for which
+capabilities of the typical fitting technology. Presently a typical
+global PDF fit includes thousands of hadronic data points for which
 predictions have to be computed thousands of times during the
@@
 \end{equation}
-where $Q^2$ is the typical hard scale of the process, the index
-$s$ sums over the active partonic subprocesses in the calculation, $p_{\rm LO}$ is the leading-order power of $\alpha_s$ for the process
-and $\hat{\sigma}^{(p)(s)}$ is the N$^p$LO cross-section for the partonic subprocess scattering $s\to X$. $F^{(s)}$ represents the 
-subprocess parton density:
+where $Q^2$ is the typical hard scale of the process, the index $s$
+sums over the active partonic subprocesses in the calculation, $p_{\rm
+  LO}$ is the leading-order power of $\alpha_s$ for the process and
+$\hat{\sigma}^{(p)(s)}$ is the N$^p$LO contribution to the
+cross-section for the partonic subprocess scattering $s\to
+X$. $F^{(s)}$ represents the subprocess parton density:
 \begin{equation}\label{eq:APPLsubproc}
@@
 \end{equation}
-where we use the shorthand
-$F^{(s)}_{\alpha\beta ,\tau} = F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.
-Using these expressions in the double-convolution of
-Eq.~(\ref{eq:hadconv}) one can finally obtain an expression for the
-desired cross-section which depends upon the subprocess PDFs only
-though a simple product:
+where we use the shorthand $F^{(s)}_{\alpha\beta ,\tau} =
+F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.  Using these expressions in the
+double convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an
+expression for the desired cross-section which depends upon the
+subprocess PDFs only through a simple product:
 
 \begin{equation} \label{eq:applconv}
-  \sigma_{pp\to X} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta,\tau} 
+  \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
   \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
 \end{equation}
-where
+where:
 \begin{equation}
@@
 computationally efficient inside a fit than the direct performance of
-the double-convolution.
+the double convolution.
 
@@
 A number of tools, $e.g.$ {\tt
-  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}
-or {\tt HOPPET}~\cite{Salam:2008qg}, are available which perform PDF
-evolution via an analogous procedure. {\tt APFEL} in particular, is
-able to provide the so-called \textit{evolution operator}. The
-evolution operator, is computed assuming that the DGLAP evolution can
-be encoded in a single operator whose action on the intial scale PDFs
-proceduces the final scale PDFs. Taking $Q_0$ as an initial scale and
-$Q_\tau$ as a final scale, on the $x$-space grid one can write:
+  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba},
+{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
+are available which perform PDF evolution via an analogous
+procedure. {\tt APFEL} in particular, is able to provide the so-called
+\textit{evolution operator}. The evolution operator, is computed
+assuming that the DGLAP evolution can be encoded in a single operator
+whose action on the intial scale PDFs proceduces the final scale
+PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
+on the $x$-space grid one can write:
 \begin{equation}\label{eq:fastPDFfinal_recalled}
@@
 
-Given the evolution operator, we may replace the (general-scale) PDFs
-used in the subprocess parton density Eq.~(\ref{eq:APPLsubproc}) with
-their equivalent expressions c.f Eq.~(\ref{eq:fastPDFfinal_recalled})
-like so:
+Given the evolution operator defined in
+Eq.~(\ref{eq:fastPDFfinal_recalled}), we may replace the
+(general-scale) PDFs used in the subprocess parton density
+Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions like so:
 \begin{equation}\label{eq:FK1}
@@
   \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
-  A^\tau_{\beta\gamma jl}
+  A^\tau_{\beta\gamma jl}\,,
 \end{equation}
@@
 \begin{equation}\label{eq:FKTable}
-  \widetilde{W}_{kl,\delta\gamma} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta} \sum_{\tau}
+  \widetilde{W}_{kl,\delta\gamma} = \sum_p\sum_{s}\sum_{\alpha,\beta} \sum_{\tau}
 \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
@@
 convolution also benefits from the typically reduced number of active
-partonic modes at the fitting scale, with the sum over flavours in
+partonic modes at the fitting scale, with the sum over flavors in
 Eq.~(\ref{eq:FK}) being limited to those directly used in the
@@
 results from the {\tt FK} calculation are systematically faster than
-those from {\tt APPLgrids}.  In particular, when comparing {\tt FK}
-AVX to {\tt APPLgrid} timings we obtain minimally a factor of ten
-improvement in speed for electroweak vector boson production and a
+those from {\tt APPLgrid}. In particular, when comparing {\tt
+  FK} AVX to {\tt APPLgrid} timings we obtain minimally a factor of
+ten improvement in speed for electroweak vector boson production and a
 maximum factor of 2000 improvement in predictions for inclusive jet
@@
 While sheer computational speed is typically the primary consideration
-when computing observables in a PDF fit, other factors such as table
-size in the filesystem and memory, along with the computational cost
-of pre-computing {\tt FK} tables must be considered. Indeed, the
+when com\-pu\-ting observables in a PDF fit, other factors such as
+table size in the filesystem and memory, along with the computational
+cost of pre-computing {\tt FK} tables must be considered. Indeed, the
 computation of the {\tt FK} table in Eq.~(\ref{eq:FKTable}) requires a
 great deal of operations which can be time consuming, particuarly in
-the case of source {\tt APPLgrids} with very high interpolation
+the case of source {\tt APPLgrid} files with very high interpolation
 precision.
@@
 time with {\tt APFELgrid}, {\tt FK} table file size and memory usage
-of the {\tt FK} tables arising from the same source {\tt APPLgrids} as
-discussed in Fig.~\ref{fig:timings}.  When examining the table
-generation time per point, we observe timings from a few milliseconds
-to 3.5 minutes per point, with differences arising from the varying
-grid densities used in the input {\tt APPLgrid} files.  In terms of
-the grid size on disk, {\tt FK} tables are typically larger than their
-corresponding {\tt APPLgrids}, primarily as the {\tt FK} file format
-is encoding in plain text for compatibility whereas {\tt APPLgrids}
-are expressed in binary as {\tt ROOT} files. However when measuring
-the in-memory resident set size used by the both convolution
-procedures, the amount of memory required by {\tt FK} tables is
-systematically less than {\tt APPLgrid} for all processes considered
-here by at least 75\%.
+of the {\tt FK} tables arising from the same source {\tt APPLgrid}
+files as discussed in Fig.~\ref{fig:timings}.  When examining the
+table generation time per point, we observe timings from a few
+milliseconds to 3.5 minutes per point, with differences arising from
+the varying grid densities used in the input {\tt APPLgrid} files.  In
+terms of the grid size on disk, {\tt FK} tables are typically larger
+than their corresponding {\tt APPLgrid} files, primarily as the {\tt
+  FK} file format is encoded in plain text for compatibility whereas
+{\tt APPLgrid} files are expressed in binary as {\tt ROOT}
+files. However when measuring the in-memory resident set size used by
+the both convolution procedures, the amount of memory required by {\tt
+  FK} tables is systematically less than {\tt APPLgrid} files for all
+processes considered here by at least 75\%.
 
@@
 makes the {\tt APFELgrid} code a valuable tool for modern PDF fits
-including large collider datasets and thus rely heavily upon fast
+including large collider datasets and that thus rely heavily upon fast
 calculations.
@@
 
-The {\tt APFELgrid} package and associated documentation are publicly available from the webpage:
+The {\tt APFELgrid} package and associated documentation are publicly
+available from the webpage:
 \begin{center}

### 2016-05-04 paper.tex
Abstract and intro edits

@@
 \begin{abstract}
-  We present a new software package designed to aid the inclusion of
-  hadronic observables into Parton Distribution Function (PDF)
+  We present a new software package designed to reduce the computational
+  burden of collider measurements in Parton Distribution Function (PDF)
   fits. The {\tt APFELgrid} package converts interpolated weight
   tables provided by {\tt APPLgrid} files into a more efficient format
-  for PDF fitting by the pre-computation and combination with PDF and
+  for PDF fitting by the combination with PDF and
   $\alpha_s$ evolution factors provided by {\tt APFEL}. This
@@
   scalar product. We demonstrate that our technique can lead to a
-  substantial speed improvement as compared to the existing interfaces
+  substantial speed improvement when compared to existing methods
   without any reduction in numerical accuracy.
@@
 
-The ability to perform fast and accurate cross-section predictions for
-measurements at the Large Hadron Collider (LHC) is a vital requirement
-for the most precise PDF determinations. LHC measurements have a
-unique capacity to shed light upon the dynamics of the proton and
-provide stringent constraints upon the proton PDF. In order to make
-full use of current and future LHC data, efficient strategies for the
-computation of predictions for these observables must be
-deployed. Software packages performing predictions for non-inclusive
-hadron-collider observables cannot be easily deployed in a PDF fit
-requiring rapid iteration due the time required to obtain accurate
-results (usually of the order of a few hours or more per data
-point). In order to overcome this limitation, the typical strategy
-adopted for fast cross-section prediction relies on the precomputation
-of the partonic hard cross sections in such a way that the numerical
-convolution with any set of PDFs can be quickly performed by means of
+Measurements at the Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
+of the proton and provide new constraints upon proton PDFs. However the inclusion of large collider datasets 
+into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
+theoretical predictions over the many iterations required in a fitting procedure. In order to make the fullest use of current 
+and future LHC results, efficient strategies for the computation of these observables must therefore be employed.
+
+The typical Monte Carlo software packages used to perform predictions for 
+hadron-collider observables cannot be easily deployed in a PDF fit due the processing time 
+required to obtain accurate results (usually of the order of a few hours or more per data
+point). To overcome such limitations, the typical strategy
+adopted for fast cross section prediction relies on the precomputation
+of the partonic hard cross sections in such a way that the standard numerical
+convolution with any set of PDFs can be reliably approximated by means of
 interpolation techniques.
 
-This interpolation strategy is implemented in the {\tt
+Such interpolation strategies are implementated in the {\tt
   APPLgrid}~\cite{Carli:2010rw} and {\tt
   FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the
-hard cross-sections, these packages rely on external codes to which
+hard cross sections, these packages rely on external codes to which
 they are interfaced by means of a suite of functions allowing for the
 filling of PDF- and $\alpha_s$-independent look-up tables of
-cross-section weights. Codes such as {\tt MCFM}~\cite{Campbell:2010ff}
+cross section weights. Monte Carlo programs such as {\tt MCFM}~\cite{Campbell:2010ff}
 and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
 {\tt APPLgrid}/{\tt FastNLO} and more recently de\-di\-ca\-ted
-interfaces to automated general-purpose Monte Carlo (MC) event
+interfaces to automated general-purpose event
 generators have been developed. The {\tt
   aMCfast}~\cite{Bertone:2014zva} and {\tt
-  MCgrid}~\cite{DelDebbio:2013kxa} codes have been specifically
-designed to extract the relevant information from the {\tt
+  MCgrid}~\cite{DelDebbio:2013kxa} codes can generate interpolation grids in {\tt APPLgrid}/{\tt FastNLO} format 
+ by extracting the relevant information from the {\tt
   MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
@@
 
-While these tools have proven to be extremely useful in the extraction
+While these tools have proven to be invaluable in the extraction
 of parton densities, the volume of experimental data made available by
 LHC collaborations for use in PDF fits is already stretching the
-capabilities of the typical fitting technology. Presently a typical
-global PDF fit includes thousands of hadronic data points for which
+capabilities of the typical fitting technology. A typical
+global PDF fit may now include thousands of hadronic data points for which
 predictions have to be computed thousands of times during the
-minimisation process. As a consequence the ``standard'' tools, $i.e.$
-{\tt APPLgrid} and {\tt FastNLO}, might not be fast enough to meet the
-needs of modern PDF fitters. For this reason a high-performance tool
-tailored specifically for PDF analyses is becoming increasingly
+minimisation process. As a consequence performing these predictions
+using the standard interpolating tools, $i.e.$
+{\tt APPLgrid} and {\tt FastNLO}, starts to become prohibitively time-consuming.
+For this reason a high-performance tool tailored specifically to the requirements of PDF analysis becomes increasingly
 important.
 
-The {\tt FastKernel} method, developed in the framework of the NNPDF
-global analyses~\cite{Ball:2010de,Ball:2014uwa}, was developed to
-address this problem. This method differs from the standard procedure
+The {\tt FastKernel} method was developed to
+address this problem in the context of the NNPDF global analyses~\cite{Ball:2010de,Ball:2014uwa}. This method differs from the standard procedure
 \`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
-amount of information that is precomputed prior to fitting such as to
-minimise the amount of operations to be performed during the
-fit. Specifically, the {\tt FastKernel} method relies on the
-precomputation of the partonic hard cross-sections and DGLAP evolution
-kernels and their combination into a single look-up table, here called
-a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction
-for a given hadronic observable can be obtained by a simple matrix
-product between the respective {\tt FK} table and the initial scale
-PDFs.
+amount of information that is precomputed prior to fitting so as to
+minimise the amount of operations required during the
+fit. More specifically, the {\tt FastKernel} method relies on the
+combination of precomputed hard cross sections with DGLAP evolution
+kernels into a single look-up table, here called
+a {\tt FastKernel} ({\tt FK}) table. In this way the prediction
+for a given hadronic observable can be obtained by performing a simple matrix
+product between the respective {\tt FK} table and PDFs evaluated directly at the fitting scale.
 
@@
   APFEL}
-package~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}.
+package~\cite{Bertone:2013vaa}.
 
-The outline of this paper is the following. In
+This paper proceeds as follows. In
 Sect.~\ref{sec:FastKernel} we present the technical details of the
-implementation of the {\tt FastKernel} method. Then, in
-Sect.~\ref{sec:benchmark} we illustrate timings and performance
-results of the {\tt APFELgrid} library and grids. Finally, in
+implementation of the {\tt FastKernel} method. This is followed in
+Sect.~\ref{sec:benchmark} by a performance benchmark of the
+{\tt APFELgrid} library and resulting {\tt FK} tables. Finally, in
 Sect.~\ref{sec:conclusion} we summarise the results discussed in this
@@
 
-Hadron collider observables are typically computed in QCD by a double
+Hadron collider observables are typically computed in QCD by means of a double
 convolution of parton densities with a hard scattering
-cross-section. Consider for example the calculation of a general
-cross-section $pp\to X$ with a set of PDFs $\{f\}$:
+cross section. Consider for example the calculation of a general
+cross section $pp\to X$ with a set of PDFs $\{f\}$:
 \begin{equation}\label{eq:hadconv}
@@
   \sum_{s}\sum_{p} \int dx_1\,dx_2\,
-  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,.
+  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,,
 \end{equation}
 where $Q^2$ is the typical hard scale of the process, the index $s$
-sums over the active partonic subprocesses in the calculation, $p_{\rm
+sums over the active partonic subprocesses in the calculation, $p$ sums over
+the perturbative orders used in the expansion, $p_{\rm
   LO}$ is the leading-order power of $\alpha_s$ for the process and
 $\hat{\sigma}^{(p)(s)}$ is the N$^p$LO contribution to the
-cross-section for the partonic subprocess scattering $s\to
+cross section for the partonic subprocess scattering $(s)\to
 X$. $F^{(s)}$ represents the subprocess parton density:
@@
 where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs
-that contribute to the $s$-th subprocess.  The central observation of
+contributing to the $s$-th subprocess.  The central observation of
 tools such as {\tt APPLgrid} and {\tt FastNLO} is that the PDF and
@@
 the two values of parton-$x$. For example one may represent the
-subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis
+subprocess PDFs and $\alpha_S$ in terms of Lagrange basis
 polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and
@@
   \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
-  \,\mathcal{I}_\beta(x_2)
+  \,\mathcal{I}_\beta(x_2),
 \end{array}
@@
 double convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an
-expression for the desired cross-section which depends upon the
+expression for the desired cross section which depends upon the
 subprocess PDFs only through a simple product:
@@
   \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
-  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
+  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,,
 \end{equation}
-where:
+where
 \begin{equation}
@@
 \end{equation}
-consists of the convolution of the hard cross-section with the
+consists of the convolution of the hard cross section with the
 interpolating polynomials. This information may be stored in a
-precomputed look-up table.  The final expression for the cross-section
-in Eq.~(\ref{eq:applconv}) therefore being considerably more
-computationally efficient inside a fit than the direct performance of
-the double convolution.
+precomputed look-up table.  The final expression for the cross section
+in Eq.~(\ref{eq:applconv}) is therefore a considerably simpler task to perform
+inside a fit than the direct evaluation of the double convolution.
 
@@
 A number of tools, $e.g.$ {\tt
-  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba},
+  APFEL},
 {\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
@@
 subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
-cross-section calculation:
+cross section calculation:
 \begin{equation}
@@
 In this section we shall examine the performance differences between
-the two expressions for cross-section prediction
+the two expressions for cross section prediction
 Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
@@
 $t\bar{t}$ total
-cross-sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
+cross sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
 double-differential Drell-Yan
-cross-sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
+cross sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
 data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions

### 2016-05-04 paper.tex
section 2 edits

@@
 where
-\begin{equation}
+\begin{equation} \label{eq:applgrid}
   W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
@@
 
-A number of tools, $e.g.$ {\tt
+A number of tools ($e.g.$ {\tt
   APFEL},
-{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
+{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay}) 
 are available which perform PDF evolution via an analogous
-procedure. {\tt APFEL} in particular, is able to provide the so-called
-\textit{evolution operator}. The evolution operator, is computed
-assuming that the DGLAP evolution can be encoded in a single operator
-whose action on the intial scale PDFs proceduces the final scale
-PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
-on the $x$-space grid one can write:
+interpolation procedure. In such a way PDFs at a general scale $Q_\tau$ may be
+expressed as a product of PDFs at some initial, fitting scale $Q_0$ and an
+\textit{evolution operator} obtained by the solution of the DGLAP equation.
 \begin{equation}\label{eq:fastPDFfinal_recalled}
@@
 \end{equation}
-where the operator $A$ is the solution of the DGLAP evolution equation
-with baundary condition:
-\begin{equation}
-A^\tau_{\alpha\beta, ik} =
-\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
-\end{equation}
-
-Given the evolution operator defined in
-Eq.~(\ref{eq:fastPDFfinal_recalled}), we may replace the
+where latin indices run over PDF flavour, greek indices run over points in an initial-scale interpolating $x$-grid and
+the evolution operator $A$ may be accessed directly in the {\tt APFEL} package. Given this operator, we may replace the
 (general-scale) PDFs used in the subprocess parton density
-Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions like so:
+Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions evaluated at the fitting scale as
 \begin{equation}\label{eq:FK1}
@@
 \end{equation}
-where the object:
+with the object
 \begin{equation}
@@
 \end{equation}
-combines the operations of subprocess density construction and PDF
-evolution. Going further it is possible to then use the expression for
+combining the operations of subprocess density construction and PDF
+evolution. Going further and using the expression for
 subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
-cross section calculation:
+cross section calculation we obtain
 \begin{equation}
@@
 analogously interpolated DGLAP evolution operators. This combination
-enables for a maximally computationally-efficient expression for the
+enables for a maximally efficient expression for the
 calculation of observables at hadron colliders under PDF variation.

### 2016-05-04 paper.tex
sec 3 and 4 pass

@@
 enables for a maximally efficient expression for the
-calculation of observables at hadron colliders under PDF variation.
+calculation of observables at hadron colliders under PDF variation, without invoking
+any additional approximation.
 
@@
 product in Eq.~(\ref{eq:applconv}) in several notable ways. Firstly
-the typical {\tt APPLgrid} or {\tt FastNLO} products require PDFs at a
-general scale $Q^2$, which implies performing PDF evolution {\it e.g.}
-Eq.~(\ref{eq:fastPDFfinal_recalled}) for every variation of the PDFs
-at the input scale during the fit. However in the {\tt FK} convolution
+the typical {\tt APPLgrid} or {\tt FastNLO} products use as input PDFs at a
+general scale, requiring that PDF evolution {\it e.g.}
+Eq.~(\ref{eq:fastPDFfinal_recalled}) be performed for every variation of the PDFs
+during the fit. In the {\tt FK} product
 this evolution is pre-cached at the stage of {\tt FK} table
-generation, requiring only initial-scale PDFs as an input at fitting
-time. This pre-caching of the evolution also removes the need for
-performing the sums over hard scale and perturbative order during the
-fit, further reducing the number of operations required. The {\tt FK}
-convolution also benefits from the typically reduced number of active
-partonic modes at the fitting scale, with the sum over flavors in
-Eq.~(\ref{eq:FK}) being limited to those directly used in the
+generation, requiring only initial-scale PDFs at the time of fitting. 
+This pre-caching of the evolution also removes the need to
+ sum over hard scale and perturbative order during the
+fit, further reducing the number of operations required. As the {\tt FK}
+product acts directly at the fitting scale, it benefits from the typically reduced number of active
+partonic modes, with the sum over flavors in
+Eq.~(\ref{eq:FK}) being limited to those directly parametrised in the
 fit. Having reduced the calculation to such a simple form, it is also
-possible to straightforwardly apply standard and widely-available
+straightforward to apply standard 
 computational tools such as multi-threading through e.g OpenMP or
 Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to
-further reduce computational cost.
+further reduce computational expense.
 
@@
 {\tt FK} table format is not suitable as a complete replacement for
-tools such as {\tt APPLgrid}. The pre-computation of the PDF evolution
+tools such as {\tt APPLgrid}. The precomputation of the PDF evolution
 necessarily means that all theory parameters such as perturbative
@@
 be exceptionally time consuming, the data representation in {\tt
-  APPLgrid} files allows for an efficient combination with varying
+  APPLgrid} files allows for an efficient (re)-combination with varying
 theory parameters.
@@
 
-In this section we shall examine the performance differences between
-the two expressions for cross section prediction
+We shall now examine the performance differences between
+the two expressions for fast interpolated cross section prediction
 Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
@@
 %
-To provide a comprehensive benchmark, we consider here a wide range of
+In order to provide a comprehensive benchmark, we consider here a wide range of
 processes including LHC and Tevatron electroweak vector boson
@@
 different configurations: AVX-OpenMP 2x (2 CPU cores), AVX, SSE3 and
-standard double precision dot product. Due to the inherent structural
+the standard double precision product. Due to the inherent structural
 differences between the {\tt FK} and {\tt APPLgrid} procedures,
@@
 significant performance improvements from using the {\tt FK}
-calculation even in the maximally compatible, non-SIMD case.
+calculation even without the use of SIMD or multi-threading.
 
@@
 files. However when measuring the in-memory resident set size used by
-the both convolution procedures, the amount of memory required by {\tt
+the two procedures, the amount of system memory used by {\tt
   FK} tables is systematically less than {\tt APPLgrid} files for all
@@
   FastKernel} method, it is possible to convert an {\tt APPLgrid}
-weight table into a derived format, referred to as {\tt FK} table,
-which includes PDF and $\alpha_s$ evolution. This procedure has been
+weight table into a derived format, referred to as an {\tt FK} table,
+including the effects of PDF and $\alpha_s$ evolution. This procedure has been
 implemented in the {\tt APFELgrid} package, supplied as a set of {\tt
@@
 initial scale interpolating $x$-grid and the incoming parton
-flavors. Additionally the simple structure of the resulting
-convolution makes {\tt FK} tables particularly suitable to make
+flavors. The simple structure of the resulting
+product makes {\tt FK} tables particularly suitable for the
 efficient use of computational tools such as {\tt SIMD} and {\tt
@@
 We have shown that in several practical examples the numerical
-evaluation of an {\tt FK} product is much faster than the
+evaluation of an {\tt FK} product is considerably faster than the
 corresponding {\tt APPLgrid} product, even in the case where neither
@@
 simple plain-text format in order to simplify the construction of user
-interfaces, and therefore typically are often larger than the
-corresponding {\tt APPLgrid} file. However we have shown that the
-in-memory resident set size occupied by {\tt FK} tables is typically
-smaller than that required by {\tt APPLgrid} by at least 75\%.
+interfaces, and therefore are typically larger than 
+corresponding {\tt APPLgrids}. However we have shown that the
+in-memory resident set sizes occupied by {\tt FK} tables are typically
+smaller than those required by {\tt APPLgrids}, in our examples by at least 75\%.
 
 The substantial speed improvement of {\tt FK} tables with respect to
-{\tt APPLgrid} in association with the reduction in memory footprint,
+{\tt APPLgrid} in association with the reduction in memory footprint
 makes the {\tt APFELgrid} code a valuable tool for modern PDF fits
-including large collider datasets and that thus rely heavily upon fast
-calculations.
-
+including large collider datasets.
 %We stress that {\tt APFELgrid} is not

### 2016-05-04 paper.tex
ack
fixing some typos

@@
 
-Such interpolation strategies are implementated in the {\tt
+Such interpolation strategies are implemented in the {\tt
   APPLgrid}~\cite{Carli:2010rw} and {\tt
@@
 over the initial scale interpolating $x$-grid and the incoming parton
-flavors:
+flavours:
 \begin{equation}\label{eq:FK}
@@
 product acts directly at the fitting scale, it benefits from the typically reduced number of active
-partonic modes, with the sum over flavors in
+partonic modes, with the sum over flavours in
 Eq.~(\ref{eq:FK}) being limited to those directly parametrised in the
@@
 computation of the {\tt FK} table in Eq.~(\ref{eq:FKTable}) requires a
-great deal of operations which can be time consuming, particuarly in
+great deal of operations which can be time consuming, particularly in
 the case of source {\tt APPLgrid} files with very high interpolation
@@
 initial scale interpolating $x$-grid and the incoming parton
-flavors. The simple structure of the resulting
+flavours. The simple structure of the resulting
 product makes {\tt FK} tables particularly suitable for the
@@
 
-We thanks the NNPDF Collaboration members for the support and
-motivation they gave us to write this work. V.~B. and N.~H. are
+The authors would like to thank members of the NNPDF Collaboration for their support and
+motivation for this work; particularly Juan Rojo, Luigi del Debbio and Alberto Guffanti. V.~B. and N.~H. are
 supported by an European Research Council Starting Grant ``PDF4BSM''.

### 2016-05-04 paper.tex
updating paper

@@
 those from {\tt APPLgrid}. In particular, when comparing {\tt
-  FK} AVX to {\tt APPLgrid} timings we obtain minimally a factor of
+  FK} AVX-OpenMP 2x to {\tt APPLgrid} timings we obtain minimally a factor of
 ten improvement in speed for electroweak vector boson production and a

### 2016-05-04 paper.tex
updating title

@@
 
-\title{{\tt APFELgrid}: a fast interface to hadronic observables for
-  fits of parton densities}
+\title{{\tt APFELgrid}: a high performance tool for parton density determinations}
 \tnotetext[mytitlenote]{Preprints: OUTP-16-09P, CERN-TH-2016-103}
@@
 \noindent
-{\em Manuscript Title:} {\tt APFELgrid}: a fast interface to hadronic observables for
-  fits of parton densities                                      \\
+{\em Manuscript Title:} {\tt APFELgrid}: a high performance tool for parton density determinations                                      \\
 {\em Authors:} V.~Bertone, S.~Carrazza, N.P.~Hartland                                               \\

### 2016-05-05 paper.tex
extra references

@@
 
-Measurements at the Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
-of the proton and provide new constraints upon proton PDFs. However the inclusion of large collider datasets 
+Measurements at colliders such as the Tevatron and Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
+of the proton and provide constraints upon proton PDFs~\cite{Rojo:2015acz}. However the inclusion of large hadron collider datasets 
 into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
@@
 The {\tt FastKernel} method was developed to
-address this problem in the context of the NNPDF global analyses~\cite{Ball:2010de,Ball:2014uwa}. This method differs from the standard procedure
+address this problem in the context of the NNPDF global analyses~\cite{Ball:2014uwa}. This method differs from the standard procedure
 \`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
@@
 an initial scale below the charm threshold, therefore with seven
-active partonic species.
+active partonic species. These settings are chosen so as to provide a realistic
+comparison, in a production environment the density and distribution of the
+$x$-grid may be adjusted to match interpolation requirement.
 

### 2016-05-06 paper.tex
figure mods and ttbar references

@@
 $t\bar{t}$ total
-cross sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
+cross sections~\cite{ATLAS:2012aa, Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
 double-differential Drell-Yan

### 2016-05-06 paper.tex
clarifying comments

@@
 comparison, in a production environment the density and distribution of the
-$x$-grid may be adjusted to match interpolation requirement.
+$x$-grid may be adjusted to match interpolation accuracy requirements.  For these comparisons the {\tt FK}
+table is stored as double-precision in memory for table generation and in
+single-precision for the purposes of computing the product and memory overheads.
 
@@
   FK} tables is systematically less than {\tt APPLgrid} files for all
-processes considered here by at least 75\%.
+processes considered here by at least 75\%. Note that this effect is
+in part due to the differing precisions of the default representations.
 
@@
 The authors would like to thank members of the NNPDF Collaboration for their support and
-motivation for this work; particularly Juan Rojo, Luigi del Debbio and Alberto Guffanti. V.~B. and N.~H. are
+motivation for this work; particularly Juan Rojo, Luigi del Debbio and Alberto Guffanti. We would also like
+to thank Mark Sutton for helpful comments on the paper. V.~B. and N.~H. are
 supported by an European Research Council Starting Grant ``PDF4BSM''.

### 2016-05-06 paper.tex
minor edits

@@
   We present a new software package designed to reduce the computational
-  burden of collider measurements in Parton Distribution Function (PDF)
+  burden of hadron collider measurements in Parton Distribution Function (PDF)
   fits. The {\tt APFELgrid} package converts interpolated weight
@@
 Measurements at colliders such as the Tevatron and Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
-of the proton and provide constraints upon proton PDFs~\cite{Rojo:2015acz}. However the inclusion of large hadron collider datasets 
-into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
+of the proton and provide constraints upon proton PDFs~\cite{Rojo:2015acz}. However including large hadron collider datasets 
+in PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
 theoretical predictions over the many iterations required in a fitting procedure. In order to make the fullest use of current 
@@
 The typical Monte Carlo software packages used to perform predictions for 
-hadron-collider observables cannot be easily deployed in a PDF fit due the processing time 
+hadron collider observables cannot be easily deployed in a PDF fit due the processing time 
 required to obtain accurate results (usually of the order of a few hours or more per data
@@
 LHC collaborations for use in PDF fits is already stretching the
-capabilities of the typical fitting technology. A typical
+capabilities of the typical fitting technology. A standard
 global PDF fit may now include thousands of hadronic data points for which
 predictions have to be computed thousands of times during the
-minimisation process. As a consequence performing these predictions
+minimisation process. As a consequence, performing these predictions
 using the standard interpolating tools, $i.e.$
@@
 interpolation procedure. In such a way PDFs at a general scale $Q_\tau$ may be
-expressed as a product of PDFs at some initial, fitting scale $Q_0$ and an
+expressed as a product of PDFs at some initial fitting scale $Q_0$ and an
 \textit{evolution operator} obtained by the solution of the DGLAP equation.
@@
 computational tools such as multi-threading through e.g OpenMP or
-Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to
+Single Instruction Multiple Data (SIMD) operations such as SSE or AVX to
 further reduce computational expense.
@@
 source {\tt APPLgrid} files corresponding to the majority of available
-LHC and Tevatron datasets of applicability to PDF determination.
+LHC and Tevatron datasets applicable to PDF determination.
 While the source {\tt APPLgrid} files have varying grid densities in
@@
 table is stored as double-precision in memory for table generation and in
-single-precision for the purposes of computing the product and memory overheads.
+single-precision for the purposes of computing the {\tt FK} product.
 

### 2016-05-06 paper.tex
FK titles in plots

@@
   \includegraphics[scale=0.6]{plots/t0a}
-\caption{\small Performance comparisons between {\tt APFELgrid} with
+\caption{\small Performance comparisons between {\tt FK} with
   AVX-OpenMP 2x, AVX, SSE3, double precision convolution and {\tt
